### 2019-04-17 biotime.Rmd
Search for common names with the appropriate column, give preference to species that match on the scientific column. Modify the unresolved function to gropu by the appropriate id (when do different columns are used in by_, sort will not be the appropriate unique grouping column.)

---
title: "BioTime"
author: "Kari Norman"
date: "4/1/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(here)
library(tidyverse)
library(lazyeval)
library(rdataretriever)

```

Import occurence data from Biotime
```{r}
#rdataretriever::install("biotime", connection = "csv", data_dir = here("data"))

data <- read_csv(here("data", "biotime_metadata.csv")) %>% filter(taxa %in% c("Birds", "Mammals")) %>%
  select(study_id, taxa) %>%
  left_join(read_csv(here("data", "biotime_query.csv")), by = "study_id")

```

Get mammal and trait data
```{r}
bird_trait <- read_csv(here("data/elton_traits/elton_traits_BirdFuncDat.csv"))
mamm_trait <- read.csv(here("data/elton_traits/elton_traits_MammFuncDat.csv"))
```

Check which authority would result in the most matches 
```{r}
library(taxadb)

get_match_counts <- function(data, species_col){
  species_col <- lazyeval::as_name(species_col)
  data %>% 
    select(!!species_col) %>% 
    drop_na() %>%
    distinct() %>%
    mutate(
      gbif = get_ids(!!species_col, "gbif"),
      col = get_ids(!!species_col, "col" ),
      itis = get_ids(!!species_col, "itis"),
      ncbi = get_ids(!!species_col, "ncbi"),
      wd = get_ids(!!species_col, "wd"  ),
      iucn = get_ids(!!species_col, "iucn"),
      ott = get_ids(!!species_col, "ott" ) 
    ) %>%
    #select(!!call("-", species_col)) %>%
    purrr::map_dbl(function(x) sum(!is.na(x)))
}

##Counts show that OTT is the best authority for both data sets
get_match_counts(data %>% filter(taxa == "Birds"), "genus_species")
get_match_counts(data, "genus_species")
get_match_counts(bird_trait, "scientific")
get_match_counts(mamm_trait, "scientific")
```

OTT is the best authority but doesn't have common names, so we use the next best ITIS. We'll start with BioTime data:

##Biotime

```{r}
#Getting matches by scientific name 
biotime <- data %>% 
  #filter(taxa == "Birds") %>%
  select(genus_species) %>%
  drop_na() %>%
  distinct() %>%
  mutate(id = get_ids(genus_species, "itis")) %>%
  left_join(data, by = "genus_species")

#Getting matches by sci name and common name 
biotime_ids <- by_name(unique(data$genus_species), "itis") %>%
  bind_rows(by_common(unique(data$genus_species), "itis")) %>%
  filter(taxonRank == "species") %>% #only want ID's to species, since that's the level of the trait data
  drop_na(acceptedNameUsageID)
```

Some names match to multiple acceptedUsage ID's, so we have to manually choose which one we want. In this case it's just the Blue-winged warbler
```{r}

get_dupe_ids <- function(data, orig_id){
  orig_id <- lazyeval::as_name(orig_id)

  data %>%
    distinct(acceptedNameUsageID, input, !!orig_id) %>%
    group_by(!!orig_id) %>% 
    filter(n()>1) %>% 
    ungroup()
}

unres <- get_dupe_ids(biotime_ids, "sort")
unres 
```

Add the unresolved ID to all the other ids and join with biotime data
```{r}
biotime_ids <- unres %>%
left_join(biotime_ids) %>% #join ID's back with all the columns
  filter(taxonomicStatus == "accepted") %>% #want the accepted ID from the two Warbler options
  bind_rows(biotime_ids %>% filter(!sort %in% unres$sort)) #join to resolved ID's, excluding the observations we resolved manually

biotime_data <- biotime_ids %>%
  select(id = acceptedNameUsageID,input, scientificName) %>%
  distinct() %>%
  right_join(data %>% select(-species, -genus), by = c("input" = "genus_species")) %>%
  rename(originalSciName = input)
```

##Elton Traits
Get id's for the Elton Bird trait data
```{r}
sci_names <-  by_name(unique(bird_trait$scientific), "itis") %>%
  drop_na(acceptedNameUsageID) %>%
  #get original specid so we can figure out how many don't have matches
  left_join(bird_trait %>% select(specid, scientific), by = c("input" = "scientific")) %>%
  mutate(match_type = "scientific")

com_names <- bird_trait %>% 
  filter(!specid %in% sci_names$specid) %>%
  pull(english) %>%
  unique() %>%
  by_common("itis") %>%
  #mutate(match_type = "common")
  #filter(taxonRank == "species") %>% #only want ID's to species, since that's the level of the trait data
  drop_na(acceptedNameUsageID) %>%
  left_join(bird_trait %>% select(specid, english), by = c("input" = "english"))%>%
  mutate(match_type = "common")

#there are both sci and common names, only want common name 
bird_ids <- bind_rows(sci_names, com_names)
```

Are there any unresolved ids? Yes - just one, which is a synonym to two different accepted id's
```{r}
unres_trait <- get_dupe_ids(bird_ids, "specid") %>%
  left_join(bird_ids)

unres_trait

#resolve to the desired ID (in this case the one with complete hierarchy data)
bird_ids <- unres_trait %>% 
  filter(!is.na(kingdom)) %>%
  bind_rows(bird_ids %>% filter(!sort %in% unres_trait$sort))

#only want one column for old name and one column for new name, these are the old ID columns to remove
col_ex <- c("specid", "passnonpass", "iocorder", "blfamilylatin", "blfamilyenglish", "blfamsequid", "taxo")

bird_trait_data <- bird_ids  %>%
  select(id = acceptedNameUsageID, input) %>%
  distinct() %>%
  right_join(bird_trait %>% select(.dots = -col_ex), by = c("input" = "scientific")) %>%
  rename(originalSciName = input)
```

check if all bird biotime data w/id's has an elton match. There are 54 species without a match
```{r}
biotime_data %>%
  filter(taxa == "Birds") %>%
  select(id) %>%
  distinct() %>%
  filter(!id %in% bird_trait_ids$acceptedNameUsageID) %>%
  dim()
```

and 605 species without an ID
```{r}
biotime_data %>%
  filter(taxa == "Birds") %>%
  select(id, genus_species) %>%
  distinct() %>%
  filter(is.na(id)) %>%
  dim()
```

Are there any studies for which we have trait data for all species?
```{r}
#join trait and biotime data
bird_data <- biotime_data %>%
  #filter(taxa == "Birds") %>%
  rowid_to_column("row") %>%
  left_join(bird_trait_data %>% 
              select(), 
            na_matches = "never")

coverage <- bird_data %>%
  group_by(study_id) %>%
  summarise(perc = sum(!is.na(id))/n()) %>%
  arrange(desc(perc))

head(coverage)
```