### 2019-05-20 analysis/coverage_check.Rmd
Split biotime and elton id labeling into two separate scripts, move coverage checking into analysis folder. Save all the data products in appropriate places.

---
title: "Check Coverage"
author: "Kari Norman"
date: "5/20/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
load(here::here("data", "biotime_data.rda"))
load(here::here("data", "elton_mamm.rda"))
load(here::here("data", "elton_bird.rda"))
```

#Checking Coverage of ID's and trait data -> biotime

check if all bird & mammal biotime data w/id's has an elton match. There are 258 species without a match (38 are birds).
```{r}
biotime_data %>%
  #filter(taxa == "Birds") %>%
  select(id) %>%
  distinct() %>%
  filter(!id %in% elton_bird$id) %>%
  dim()
```

and 605 bird species without an ID, 657 total species
```{r}
biotime_data %>%
  #filter(taxa == "Birds") %>%
  select(id, matchingName) %>%
  distinct() %>%
  filter(is.na(id)) %>%
  dim()
```

Are there any studies for which we have trait data for all species?
```{r}
#join trait and biotime data
bird_data <- biotime_data %>%
  #filter(taxa == "Birds") %>%
  left_join(elton_bird, by = "id",
            na_matches = "never")

coverage <- bird_data %>%
  group_by(study_id) %>%
  summarise(perc = sum(!is.na(specid))/n()) %>%
  arrange(desc(perc))

head(coverage)
```