### 2019-07-02 data-raw/elton.Rmd
Include all trait data, even species without id's. Standardize origin sci name column name.

---
title: "elton trait"
author: "Kari Norman"
date: "5/20/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(here)
library(tidyverse)
library(lazyeval)
library(rdataretriever)
library(taxadb)

source(here::here("data-raw", "helper_functions.R"))
```

Get mammal and trait data
```{r}
bird_trait <- read_csv(system.file("extdata", "elton_traits/elton_traits_BirdFuncDat.csv", package = "biodivTS"))
mamm_trait <- read.csv(system.file("extdata", "elton_traits/elton_traits_MammFuncDat.csv", package = "biodivTS"))
```

Check which authority would result in the most matches 
```{r}
##Counts show that OTT is the best authority for both data sets
get_match_counts(bird_trait, "scientific")
get_match_counts(mamm_trait, "scientific")
```

Get id's for the Elton trait data
```{r}
get_trait_ids <- function(data, provider, id, common = TRUE){
  #id <- lazyeval::as_name(id)
  sci_match <-  by_name(unique(data$scientific), provider) %>%
    drop_na(acceptedNameUsageID) %>%
    #get original specid so we can figure out ho w many don't have matches
    left_join(data %>% select_(id, "scientific"), by = c("input" = "scientific")) %>%
    mutate(match_type = "scientific")

  if(common){
    com_names <- data %>% 
      #the id argument should be inplace of specid, but it doesn't work for some reason
      filter(!specid %in% sci_match$specid) %>%
      pull(english) %>%
      unique() %>%
      by_common("itis") %>%
      #mutate(match_type = "common")
      #filter(taxonRank == "species") %>% #only want ID's to species, since that's the level of the trait data
      drop_na(acceptedNameUsageID) %>%
      left_join(bird_trait %>% select_(id, "english"), by = c("input" = "english"))%>%
      mutate(match_type = "common")

    #there are both sci and common names, only want common name 
    return(bind_rows(sci_match, com_names))
  }else{ return(sci_match)}
}

bird_ids <- get_trait_ids(bird_trait, "itis", "specid")
mammal_ids <- get_trait_ids(mamm_trait, "itis", "msw3_id", FALSE)

```

Are there any unresolved ids for birds? Yes - just one, which is a synonym to two different accepted id's
```{r}
unres_trait <- get_dupe_ids(bird_ids, "specid") %>%
  left_join(bird_ids)

unres_trait

#resolve to the desired ID (in this case the one with complete hierarchy data)
bird_ids <- unres_trait %>% 
  filter(!is.na(kingdom)) %>%
  bind_rows(bird_ids %>% filter(!sort %in% unres_trait$sort))
```

What about for mammals? Yes - also a synonym to two different accepted id's. Further research shows that the Callospermophilus saturatus uses the subgenus, so we'll keep the name with the genus 
```{r}
unres_mamm_trait <- get_dupe_ids(mammal_ids, "msw3_id") %>% 
  left_join(mammal_ids)

unres_mamm_trait

mammal_ids <- unres_mamm_trait %>% 
  filter(acceptedNameUsageID == "ITIS:632452") %>%
  bind_rows(mammal_ids %>% filter(!sort %in% unres_mamm_trait$sort))
```

Join bird ids back to original trait data
```{r}
#only want one column for old name and one column for new name, these are the old ID columns to remove
col_ex <- c("passnonpass", "iocorder", "blfamilylatin", "blfamilyenglish", "blfamsequid", "taxo")

#join on scientific and common
sci_matches <- bird_ids  %>%
  filter(match_type == "scientific") %>%
  select(id = acceptedNameUsageID, input, scientificName) %>%
  distinct() %>%
  right_join(bird_trait %>% select(.dots = -col_ex), by = c("input" = "scientific"), na_matches = "never") %>%
  #drop_na(id) %>%
  distinct() %>%
  select(-english)

comm_matches <- bird_ids  %>%
  filter(match_type == "common") %>%
  select(id = acceptedNameUsageID, input, scientificName) %>%
  distinct() %>%
  right_join(bird_trait %>% filter(!specid %in% sci_matches$specid) %>% select(.dots = -col_ex), by = c("input" = "english"), na_matches = "never") %>%
  distinct() %>%
  rename(matchingName = input) %>%
  select(-scientific)

#there are more rows after ID data is joined because for species that matched on common names there may be more than one scientificName 
elton_bird <- bind_rows(sci_matches, comm_matches) %>% rename(sourceName = input)
```

Do the same for mammal trait data
```{r}
elton_mamm <- mammal_ids %>%
  select(id = acceptedNameUsageID, input, scientificName) %>%
  right_join(mamm_trait %>% select(-mswfamilylatin), by = c("input" = "scientific"), na_matches = "never") %>%
  distinct() %>%
  rename(sourceName = input)
```

```{r}
usethis::use_data(elton_bird, elton_mamm)
```