### 2019-07-02 data-raw/amphibio.Rmd
Standardize name of the original sci name column.

---
title: "AmphibBio"
author: "Kari Norman"
date: "5/3/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)

source(here::here("data-raw", "helper_functions.R"))
```

```{r}
#original data from https://figshare.com/articles/Oliveira_et_al_AmphiBIO_v1/4644424

amph <- read_csv(system.file("extdata", "amphibio/AmphiBIO_v1.csv", package = "biodivTS"))

amph_ids <- amph %>%
  mutate(id = get_ids(Species, "itis")) %>%
  filter(!is.na(id)) %>%
  mutate(scientificName = get_names(id, "itis")) %>%
  select(-c("Order", "Family", "Genus")) %>%
  rename(sourceName = Species)

# Using by_ doesn't give any additional ids
# amph_ids <- by_name(amph$Species, "itis") %>%
#   filter(taxonRank == "species") %>% #only want ID's to species, since that's the level of the trait data
#   drop_na(acceptedNameUsageID)

```