### 2019-09-03 data-raw/web_scrape.Rmd
Web scraping fishtraits database with rselenium.

---
title: "fishtraits"
author: "Kari Norman"
date: "8/28/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(RSelenium)
```

```{r}
clickObject <- function(selector){
  object <- rd$findElement(using = 'css selector', selector)
  object$clickElement()
}
```
connect to Selenium server and navigate to the page
```{r}
driver <- rsDriver(browser = c("chrome"), chromever = "76.0.3809.126")
rd <- driver[["client"]]
rd$open()

rd$navigate("http://www.fishtraits.info/search/taxo/")
```

see if we can get the list of all possible search items for the FID dropdown (second is dependent on the first input)
```{r}
#click on drop down element
clickObject('.ui-combobox .ui-icon-triangle-1-s')

#get text of the list
fid_list <- rd$findElement('css selector', '#ui-id-1')
fid_text <- unlist(strsplit(unlist(fid_list$getElementText()), "\n"))
```

Choose FID and Genus
```{r}
#select and search the FID dropdown
fid <- rd$findElement(using = 'css selector', "#fid-combo")
fid$clearElement()
#pass to search box
fid$sendKeysToElement(list(fid_text[1]))
#click matching one
clickObject("#ui-id-1")

#get list of Genus's that are now possible
clickObject(".ui-combobox2 .ui-icon-triangle-1-s")
genus_list <- rd$findElement('css selector', '#ui-id-2')
genus_text <- unlist(strsplit(unlist(genus_list$getElementText()), "\n"))

#select and search the Genus dropdown
genus <- rd$findElement(using = 'css selector', "#genus-combo")
genus$clearElement()
genus$sendKeysToElement(list(genus_text[1]))
#click matching one
clickObject('#ui-id-2')
```

Pick the columns we want
```{r}
#make sure the appropriate column boxes are checked
columns <- rd$findElement(using = 'css selector', '#ui-accordion-ac1-header-1')
columns$clickElement()

#selectors for all boxes that need to be checked
```

Submit search and download
```{r}
#select search box
clickObject("#searchSubmit")
#and download
clickObject("#downloadbtn")
```

```{r}
driver$server$stop()
```