### 2019-09-03 data-raw/web_scrape.Rmd
loop across FID's and genus's

---
title: "fishtraits"
author: "Kari Norman"
date: "8/28/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(RSelenium)
```

```{r}
clickObject <- function(selector){
  object <- rd$findElement(using = 'css selector', selector)
  object$clickElement()
}
```
connect to Selenium server and navigate to the page
```{r}
driver <- rsDriver(browser = c("chrome"), chromever = "76.0.3809.126")
rd <- driver[["client"]]
rd$open()

rd$navigate("http://www.fishtraits.info/search/taxo/")
```

see if we can get the list of all possible search items for the FID dropdown (second is dependent on the first input)
```{r}
#click on drop down element
clickObject('.ui-combobox .ui-icon-triangle-1-s')

#get text of the list
fid_list <- rd$findElement('css selector', '#ui-id-1')
fid_text <- unlist(strsplit(unlist(fid_list$getElementText()), "\n"))
```

Choose FID and Genus
```{r}

for (i in 1:length(fid_text)) {
  #select and search the FID dropdown
  fid <- rd$findElement(using = 'css selector', "#fid-combo")
  fid$clearElement()
  #pass to search box
  fid$sendKeysToElement(list(fid_text[i]))
  #click matching one
  clickObject("#ui-id-1")

  #get list of Genus's that are now possible
  clickObject(".ui-combobox2 .ui-icon-triangle-1-s")
  genus_list <- rd$findElement('css selector', '#ui-id-2')
  genus_text <- unlist(strsplit(unlist(genus_list$getElementText()), "\n"))

  for (j in 1:length(genus_text)) {
    #select and search the Genus dropdown
    genus <- rd$findElement(using = 'css selector', "#genus-combo")
    genus$clearElement()
    genus$sendKeysToElement(list(genus_text[j]))
    #click matching one
    clickObject('#ui-id-2')

    #make sure the appropriate column boxes are checked
    columns <- rd$findElement(using = 'css selector', '#ui-accordion-ac1-header-1')
    columns$clickElement()

    #selectors for all boxes that need to be checked

    #select search box
    clickObject("#searchSubmit")
    #and download
    clickObject("#downloadbtn")
  }
}
```

Pick the columns we want
```{r}

```

Submit search and download
```{r}

```

```{r}
driver$server$stop()
```