### 2019-09-19 data-raw/fishbase.Rmd
Get traits using rfishbase

---
title: "Fishbase"
author: "Kari Norman"
date: "9/19/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(rfishbase)

```

Potential tables of interest: diet, ecology, ecosystem, fecundity, fooditems, maturity, popchar, popgrowth, reproduc, spawning, speed

other: morphdat, oxygen

```{r}
docs(table = "diet")
```