### 2019-10-03 data-raw/amphibio.Rmd

---
title: "AmphibBio"
author: "Kari Norman"
date: "5/3/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(taxadb)

source(here::here("data-raw", "helper_functions.R"))
```

Get id's for amphibio database from itis
```{r}
#original data from https://figshare.com/articles/Oliveira_et_al_AmphiBIO_v1/4644424

amph <- read_csv(system.file("extdata", "amphibio/AmphiBIO_v1.csv", package = "biodivTS")) %>%
  rename_all(tolower) %>%
  rename(diet_leaves = leaves, diet_flowers = flowers, diet_seed = seeds, diet_fruit = fruits, diet_inv = arthro, diet_vert = vert,
         diurnal = diu, nocturnal = noc, crepuscular = crepu, bodymass_value = body_mass_g) %>%
  mutate(diet_seed = as.numeric(diet_seed))

amphibio <- amph %>%
  mutate(id = get_ids(species, "itis")) %>%
  filter(!is.na(id)) %>%
  mutate(scientificName = get_names(id, "itis")) %>%
  select(id, scientificName, species) %>%
  right_join(select(amph, -id), by = "species") %>%
  rename(sourceName = species) %>%
  select(-c("order", "family", "genus"))

# Using by_ doesn't give any additional ids
# amphibio <- by_name(amph$Species, "itis") %>%
#   filter(taxonRank == "species") %>% #only want ID's to species, since that's the level of the trait data
#   drop_na(acceptedNameUsageID)
```

Check for matches from other providers
```{r}
syn_res <- match_providers(amphibio, "itis")

#names that resolve to more than one id
alt_dupes <- syn_res %>%
  select(acceptedNameUsageID, input) %>%
  distinct() %>%
  group_by(input) %>%
  filter(n() > 1)

add_syn_res <- syn_res %>% 
  filter(!input %in% alt_dupes$input) %>% #excluding duplicate matched species
  select(id = acceptedNameUsageID, sourceName = input, scientificName) %>%
  distinct() %>%
  group_by(sourceName) %>%
  top_n(1, scientificName) %>%
  left_join(amphibio %>% select(-c(id, scientificName)), by = "sourceName", na_matches =  "never") %>%
  ungroup() %>%
  distinct()

amphibio <- amphibio %>% 
  filter(!sourceName %in% add_syn_res$sourceName) %>%
  bind_rows(add_syn_res)
```

Select relevant trait columns
```{r}
amphibio <- amphibio %>%
  select(-obs)
```

Check for multiple matches to the same ID
```{r}
amphibio <- undupe_ids(amphibio)
```

```{r}
usethis::use_data(amphibio)
```