### 2020-09-08 analysis/merge_data.Rmd
Add number of plots to each study.

---
title: "Check Coverage"
author: "Kari Norman"
date: "5/20/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(FD)
devtools::load_all()
```

```{r}
load(here::here("data", "biotime_data.rda"))
load(here::here("data", "elton_mamm.rda"))
load(here::here("data", "elton_bird.rda"))
load(here::here("data", "amphibio.rda"))
```

#Checking Coverage of ID's and trait data -> biotime

check if all bird & mammal biotime data w/id's has an elton match. There are 18 species without a match.
```{r}
no_trait <- biotime_data %>%
  filter(taxa %in% c("Birds", "Mammals")) %>%
  select(id, sourceName, taxa) %>%
  drop_na(id) %>%
  distinct() %>%
  filter(!id %in% elton_bird$id, !id %in% elton_mamm$id)

dim(no_trait)
```

and 25268 species without an ID, some of which are species codes
```{r}
biotime_data %>%
  #filter(taxa == "Birds") %>%
  select(id, sourceName, taxa) %>%
  filter(is.na(id)) %>%
  distinct() %>%
  dim()
```

Are there any studies for which we have trait data for all species?
```{r}
merge_data <- function(trait_data){
  biotime_data %>%
    select(-scientificName) %>%
    rename(biotimeName = sourceName) %>%
  left_join(trait_data %>% rename(traitName = sourceName), by = "id",
            na_matches = "never") #%>%
   # drop_na(scientificName)
}
```

Get data w/traits for all studies that have >85% trait coverage

`id` is `NA` if there was no `id` match
`biotimeName` or `traitName` is `NA` if that ID is the result of an average of multliple species matches
`scientificName` is `NA` when there was no trait data for a given ID
```{r}
#set coverage threshold for analysis dataset
percent <- 0.85

traits <- list(elton_bird, elton_mamm, amphibio)

comp_biotime_traits <- map_dfr(traits, merge_data) %>%
  right_join(biotime_data %>% select(-sourceName, -scientificName)) %>%
  distinct()

#always check that one species didn't match to multiple datasets, below should return and empty dataframe
#comp_biotime_traits %>% group_by(record_id) %>% filter(n() > 1) %>% dim()

#this gives a coverage for each trait dataset, so potentially a lot of coverage with one dataset and no coverage on all the rest
#need to get rid of zeros when there is coverage from another dataset so we only have one entry per study
coverage <- comp_biotime_traits %>%
  select(study_id, scientificName) %>%
  distinct() %>%
  group_by(study_id) %>%
  summarise(perc = sum(!is.na(id))/n()) %>%
  arrange(desc(perc))

inc_coverage <- coverage %>% filter(perc > percent)

biotime_traits <- comp_biotime_traits %>%
  #get studies with appropriate coverage and drop observations that we don't have trait data for
  filter(study_id %in% inc_coverage$study_id, !is.na(scientificName)) %>% 
  mutate(plot_id = paste0(study_id, plot))
```

```{r}
usethis::use_data(biotime_traits)
pins::pin(biotime_traits, board = "github")
```

Create table of info for each study
```{r}
study_table <- biotime_data %>%
  group_by(study_id) %>%
  mutate(min_year = min(year), max_year = max(year), n_year = n_distinct(year), plots = n_distinct(plot)) %>% 
  select(study_id, taxa, min_year, max_year, plots) %>%
  distinct() %>%
  left_join(coverage %>% select(study_id, coverage = perc), by = "study_id")

usethis::use_data(study_table)
pins::pin(study_table, board = "github")
```