### 2020-11-20 analysis/visualization.Rmd
Modify column names to match new dataframe.

---
title: "Visualization"
author: "Kari Norman"
date: "9/18/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

```{r}
metrics <- pins::pin_get("rarefied-medians", board = "github") %>%
  pivot_longer(., cols = setdiff(everything(), one_of("rarefyID", "STUDY_ID", "SamplePool", "SampleN", "num_years", "duration", "startYear", "endYear", "YEAR", "cell", "rarefied", "type", "realm", "climate", "habitat", "biome_map", "taxa", "organisms", "cent_lat", "cent_long", "abundance_type", "BROAD_TYPE", "cYEAR", "taxa_mod", "climate_mod")), names_to = "metric", values_to = "value") %>%
  rename_with(tolower, .cols = setdiff(everything(), one_of("rarefyID"))) 
```

```{r}
usa <- map_data("usa")

plt <- ggplot() +
  geom_polygon(data = usa, aes(x=long, y = lat, group = group), fill = NA, color = "black") +
  coord_fixed(1.3)

plt +
  geom_point(data = biotime_data %>% filter(taxa == "Fish"), aes(x = longitude, y = latitude), color = "black", size = .25)

```

#Visualizing FD trends
```{r}
#load(here::here("data", "fd_table.rda"))
fd_table <- pins::pin_get("fd-table", board = "github")

fric_plot <- fd_table %>%
  filter(metric == "FRic") %>%
  ggplot(aes(
    x = year,
    y = value,
    colour = as.factor(plot)
  )) +
  geom_line() +
  scale_x_continuous(breaks = scales::pretty_breaks()) +
  facet_wrap( ~ as.factor(plot_id), scales = "free") +
  theme(legend.position = "none") +
  ylab("Fric")

ggsave("fric_trends.png", fric_plot)
```

Reproduce Dornelas figure
```{r}
#get climate classification
metadata <- read_csv(system.file("extdata", "biotime/biotime_metadata.csv", package = "biodivTS"))

fric_trend <- fd_table %>%
  filter(metric == "FRic") %>%
  left_join(metadata %>% select(study_id, climate)) %>%
  ggplot(aes(x = year, y = log(value))) +
  # geom_smooth(aes(color = climate, group = plot_id), method = "glm", se = FALSE) +
  # geom_point(aes(color = climate, group = plot_id)) +
  geom_smooth(aes(group = plot_id), color = "grey", method = "glm", se = FALSE) +
  geom_point(aes(group = plot_id) ,color = "grey") +
  geom_smooth(color = "black", method = "glm", se = FALSE) +
  theme(legend.position = "none") 
ggsave("fric_trends.png", fric_plot)
```

metric plotting function
```{r}
library(lme4)

plot_metric <- function(metric_name, ylabel){
  data <- metrics %>%
    filter(metric == metric_name, !is.na(value), !is.infinite(value)) %>%
    mutate(logvalue = log(value))

  mod <- lmer(logvalue ~ year + (1|rarefyID), data = data)
  coefs <- summary(mod)$coefficients

  print(metric_name) 
  print(summary(mod))
  confint(mod)

  pred <- data.frame(year = min(data$year):max(data$year)) %>%
    mutate(pred = coefs[1,1] + year * coefs[2,1])

  trend_plot <- data %>%
    left_join(pred, by = "year") %>%
    #left_join(metadata %>% select(study_id, climate)) %>%
    ggplot(aes(x = year, y = logvalue)) +
    # geom_smooth(aes(color = climate, group = rarefyID), method = "glm", se = FALSE) +
    # geom_point(aes(color = climate, group = rarefyID)) +
    geom_point(aes(group = rarefyID) ,color = "grey") +
    geom_smooth(aes(group = rarefyID), color = "darkgrey", method = "glm", se = FALSE) +
    #geom_smooth(color = "black", method = "glm", se = FALSE) +
    geom_line(mapping = aes(y = pred)) +
    theme(legend.position = "none") +
    theme_classic() +
    ylab(ylabel)

  ggsave(paste0(metric_name, "_trend.png"), trend_plot)
  return(trend_plot)
}

plot_metric("FRic", "log(FD)")

plot_map <- tibble(metric_name = c("FRic", "FEve", "FDiv", "FDis"), 
           ylabel = c("log(FD)", "log(Evenness)", "log(Divergence)", "log(Dispersion)"))
pmap(plot_map, plot_metric)
```

get model info
```{r}
get_metric_conint <- function(metric_name){
 data <- metrics %>%
    filter(metric == metric_name) %>%
    mutate(rarefyID = as.factor(rarefyID), logValue = log(value))

  mod <- lmer(logValue ~ year + (1|rarefyID), data = data)
  coefs <- summary(mod)$coefficients

  confint(mod)
  # as_tibble(confint(mod)) %>%
  #   mutate(metric = metric_name)
}

map(c("FRic", "FEve", "FDiv", "FDis"), get_metric_conint)
```