### 2021-03-11 analysis/01_filter_by_traits.Rmd
allow 75% trait coverage

---
title: "Check Coverage"
author: "Kari Norman"
date: "5/20/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(FD)
devtools::load_all()
```

```{r}
pins::board_register_github(name = "github", repo = "karinorman/biodivTS_data", branch = "master")

biotime_data <- pins::pin_get("biotime-data", board = "github")
elton_mamm <- pins::pin_get("elton-mamm", board = "github")
elton_bird <- pins::pin_get("elton-bird", board = "github")
amphibio <- pins::pin_get("amphibio", board = "github")

biotime_data <- biotime_data %>%
  ungroup() %>%
  mutate(obsID = row_number())
```

#Checking Coverage of ID's and trait data -> biotime

check if all bird & mammal biotime data w/id's has an elton match. There are 18 species without a match.
```{r}
no_trait <- biotime_data %>%
  filter(taxa %in% c("Birds", "Mammals")) %>%
  select(id, sourceName, taxa) %>%
  drop_na(id) %>%
  distinct() %>%
  filter(!id %in% elton_bird$id, !id %in% elton_mamm$id)

dim(no_trait)
```

and 25268 species without an ID, some of which are species codes
```{r}
biotime_data %>%
  #filter(taxa == "Birds") %>%
  select(id, sourceName, taxa) %>%
  filter(is.na(id)) %>%
  distinct() %>%
  dim()
```

Are there any studies for which we have trait data for all species?
```{r}
merge_data <- function(trait_data){
  biotime_data %>%
    select(-scientificName) %>%
    rename(biotimeName = sourceName) %>%
  left_join(trait_data %>% rename(traitName = sourceName), by = "id",
            na_matches = "never") %>%
   drop_na(scientificName)
}
```

Get data w/traits for all studies that have >85% trait coverage

`id` is `NA` if there was no `id` match
`biotimeName` or `traitName` is `NA` if that ID is the result of an average of multliple species matches
`scientificName` is `NA` when there was no trait data for a given ID
```{r}
#set coverage threshold for analysis dataset
percent <- 0.75

traits <- list(elton_bird, elton_mamm, amphibio)

trait_merge <- map_dfr(traits, merge_data)

comp_biotime_traits <- trait_merge %>%
  bind_rows(biotime_data %>% 
              filter(!obsID %in% trait_merge$obsID) %>%
              select(-scientificName) %>%
              rename(biotimeName = sourceName))# %>%
  #distinct()

#always check that one species didn't match to multiple datasets, below should return and empty dataframe
#comp_biotime_traits %>% group_by(record_id) %>% filter(n() > 1) %>% dim()

coverage <- comp_biotime_traits %>%
  select(study_id, id, biotimeName, traitName) %>%
  distinct() %>%
  group_by(study_id) %>%
  summarise(perc = sum(!is.na(traitName))/n()) %>%
  arrange(desc(perc))

inc_coverage <- coverage %>% filter(perc > percent)

biotime_traits <- comp_biotime_traits %>%
  #get studies with appropriate coverage and drop observations that we don't have trait data for, and unlabeled non-id'd species
  filter(study_id %in% inc_coverage$study_id, !is.na(scientificName)) %>% 
  group_by(rarefyid, year) %>%
  filter(n_distinct(id) > 4) %>%
  ungroup()
```

Get biotime data filtered by trait coverage
```{r}
bt_cols <- colnames(biotime_data)[!colnames(biotime_data) %in% 
                                         c("sourceName", "obsID", "scientificName")]
bt_traitfiltered <- biotime_traits %>% 
  select(all_of(bt_cols)) %>%
  rename(species = id)

#save locally and in cache
usethis::use_data(bt_traitfiltered)
pins::pin(bt_traitfiltered, board = "github")
```

Create a reference table of trait data
```{r}
trait_cols <- c(colnames(amphibio), colnames(elton_bird), colnames(elton_mamm))
trait_cols <- trait_cols[!trait_cols %in% c("sourceName", "scientificName")]

trait_ref <- biotime_traits %>%
  select(all_of(trait_cols)) %>%
  rename(Species = id) %>%
  distinct()

#save locally and in cache
usethis::use_data(trait_ref)
pins::pin(trait_ref, board = "github")
```

Create table of info for each study
```{r}
metadata <- pins::pin_get("metadata", board = "github")

study_table <- biotime_traits %>%
  select(study_id, plot) %>%
  count(study_id, name = "n_plots") %>% 
  left_join(metadata %>% select(study_id, realm, climate, habitat, biome_map, taxa, 
                                organisms, start_year, end_year), by = "study_id") %>%
  left_join(coverage %>% select(study_id, coverage = perc), by = "study_id")

#save locally and in cache
usethis::use_data(study_table)
pins::pin(study_table, board = "github")
```