### 2021-03-16 analysis/rarefy_null_models.Rmd
first commit

---
title: "Rarefied Null Models"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(FD)
library(furrr)

pins::board_register_github(repo = "karinorman/biodivTS_data", branch = "master")
```

Get dataframe of filenames and corresponding rarefyID's
```{r}
path <- here::here("data", "rarefied_samples")
files <- list.files(path) %>%
  paste0(path, "/", .)

samples <- purrr::map_dfr(files, function(x) {load(x)
  return(rare_comm_save)
	}) %>% distinct()
```

Function to get all rarefied samples for an ID
```{r}
# get_rarefy_samps <- function(id, path = here::here("data", "rarefied_samples")){
#   print(id)
#   samp_data <- list.files(path = path, pattern = paste0( "*.", id, ".*")) %>% 
#     paste0(path, "/", .) %>%
#      purrr::map_dfr(., function(x) {load(x)
#        return(rare_comm_save)
#        }) 
#   return(samp_data)
# }
```
Get species pools for each timeseries
```{r}
bt_grid_filtered <- pins::pin_get("bt-traitfiltered", board = "github")
trait_ref <- pins::pin_get("trait-ref", board = "github")

species_pools <- bt_grid_filtered %>%
  select(rarefyid, species) %>% 
  distinct()

rm(bt_grid_filtered)
```

Simulation functions
```{r}
get_plot_sample <- function(file, n){
  #print(filter_id)
  load(file)
  comm <- rare_comm_save %>%
    pivot_longer(cols = starts_with("ITIS"), names_to = "species", values_to = "weight") %>%
    filter(weight != 0) %>%
    rename_with(tolower)

  #check that there are at least two species for all years, otherwise stop sampling
  species_count <- comm %>% select(year, species) %>% count(year)
  if (min(species_count$n) < 2){
    save_dir <- here::here("data", "no_null_sample")
    fs::dir_create(save_dir)
    write.csv(rare_comm_save, paste0(save_dir, "/", filter_rarefyid, "_samp", samp_rarefy_resamp, ".csv"))
  }
  filter_rarefyid <- unique(comm$rarefyid) %>% unlist()
  samp_rarefy_resamp <- unique(comm$rarefy_resamp) %>% unlist()

  # plot_data <- data %>% filter(rarefyID == filter_id) %>%
  #   select(year, id, weight)

  species_pool <- species_pools %>%
    filter(rarefyid == filter_rarefyid) %>%
    pull(species) %>%
    na.omit() %>%
    unique() #get column of unique scientific names

  get_samp_species <- function(data){
    samp_species <- sample(species_pool, n_distinct(data$species))
    samp_abun <- sample(data$weight)

    df <- as_tibble(cbind(samp_species, samp_abun)) %>% mutate(year = unique(data$year))
    return(df)
  }

  site_samples <- data.frame()
  for (j in 1:n){
    iter_samples <- data.frame()
    for(i in unique(comm$year)){
      year_data <- comm %>% filter(year == i)
      iter_samples <- rbind(iter_samples, get_samp_species(year_data))
    }
    site_samples <- rbind(site_samples, iter_samples %>% mutate(n = j))
  }
  plot_samples <- site_samples %>%
    mutate(rarefyid = filter_rarefyid, rarefy_resamp = samp_rarefy_resamp) %>%
    pivot_wider(names_from = samp_species, values_from = samp_abun)

  save_dir <- here::here("data", "null_samples")
  fs::dir_create(save_dir)
  write.csv(plot_samples, paste0(save_dir, "/", filter_rarefyid, "_samp", samp_rarefy_resamp, ".csv"))

  return(plot_samples)
}
```

Get sample dataframe
```{r}
future::plan(multiprocess)
sample_occurrence <- future_map(files, get_plot_sample, n = 100) %>%
  bind_rows() %>%
  replace(is.na(.), 0)
```
Figure out which samples didn't get null samples
```{r}
null_files <- list.files(here::here("data", "null_samples")) %>%
  paste0(here::here("data", "null_samples"), "/", .)

null_samples <- map_dfr(null_files, ~read_csv(.x) %>% select(rarefyid, rarefy_resamp) %>% distinct()) %>%
  mutate(rarefy_samp_file = paste0(path, "/count_cell", rarefyid, "_sample", rarefy_resamp, ".rda"))

missing_files <- setdiff(files, null_samples$rarefy_samp_file)

future::plan(multiprocess)
sample_occurrence <- future_map(missing_files, get_plot_sample, n = 100) %>%
  bind_rows() %>%
  replace(is.na(.), 0)
```

Get richness levels for a rarefyID
```{r}
get_rarefy_richness <- function(id){
  get_rarefy_samps(id) %>%
    pivot_longer(starts_with("ITIS:"), names_to = "speciesID", values_to = "abundance") %>% 
    select(YEAR, rarefy_resamp, speciesID) %>%
    group_by(rarefy_resamp, YEAR) %>% 
    summarise(richness = n_distinct(speciesID)) %>% 
    distinct()
}

rich_levels <- map_dfr(ids, get_rarefy_richness)
```