### 2021-04-09 analysis/invasive_data.Rmd
Function to read in invasive species data

---
title: "GBIF invasive species"
author: "Kari Norman"
date: "3/24/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(rgbif)
library(geonames)
library(countrycode)
library(taxadb)
```

Get GBIF data
```{r}
#get all the datasets from the Invasive Species Specialist Group
datasets <- dataset_search(publishingOrg = "cdef28b1-db4e-4c58-aa71-3c5238c2d0b5", limit = 300)
datasets_list <- datasets$data

occ <- occ_download(dataset = "b351a324-77c4-41c9-a909-f30f77268bc4")
```

Figure out which country each study is located in 
```{r}
pins::board_register_github(name = "github", repo = "karinorman/biodivTS_data", branch = "master")
options(geonamesUsername="karinorman")

obs_data <- pins::pin_get("bt-traitfiltered", board = "github") %>%
  select(study_id, species) %>%
  distinct()

meta <- pins::pin_get("meta", board = "github") %>%
  select(study_id, cent_lat, cent_long) %>%
  filter(study_id %in% obs_data$study_id)

get_GNcountry <- possibly(GNcountryCode, otherwise = NA_character_)

get_country <- function(study_id, cent_lat, cent_long){
  #con_data <- get_GNcountry(lat = cent_lat, lng = cent_long, lang = "en", radius = 50)
  con_data <- get_GNcountry(lat = cent_lat, lng = cent_long, radius = 50)
  name <- ifelse(length(con_data) == 1, con_data, con_data$countryName)

  return(tibble(study_id = study_id, country = name))
}

country_data <- pmap_dfr(meta, get_country) %>%
  #hand resolve study_ids in the ocean
  mutate(country = case_when(
  study_id == 108 ~ "Namibia", 
  study_id %in% c(56, 166, 169) ~ "United States", 
  study_id == 172 ~ "Morocco", 
  TRUE ~ country
)) %>%
  mutate(code = countryname(country, destination = "genc2c")) %>%
  mutate(code = case_when(
    code == "CA" ~ "US", 
    code == "PR" ~ "DO",
    TRUE ~ code
  ))

obs_data <- obs_data %>%
  left_join(country_data)
```

Read in Invasive Data
```{r}
invasive_files <- list.dirs(here::here("data", "invasive"))

invasive_data <- map_dfr(invasive_files[2:21], function(x){
  taxon <- read_tsv(paste0(x, "/", "taxon.txt"))
  profile <- read_tsv(paste0(x, "/", "speciesprofile.txt"))
  dist <- read_tsv(paste0(x, "/", "distribution.txt")) %>%
    select(id, countryCode)

  data <- left_join(taxon, profile) %>% left_join(dist)
  return(data)
})
```

Substituting years:
- orig: Canada, data: United States
- orig: Puerto Rico, data: Dominican Republic

Get datasets relevant for observed studies
```{r}
countries <- unique(country_data$country)
datasets_list %>% filter(str_detect(datasetTitle, countries)) %>% View()

```