### 2021-04-13 analysis/func_distinctiveness.Rmd
First commit

---
title: "Functional Rarity"
output: html_document
---
```{r}
library(funrar)
library(biodivTS)
```

Calculate functional distinctiveness for each species based on the species pool (defined by all species observed over the course of the timeseries)

Get data
```{r}
pins::board_register_github(name = "github", repo = "karinorman/biodivTS_data", branch = "master")

##	Get the gridded data locally
bt_data <- pins::pin_get("bt-traitfiltered", board = "github") %>%
  rename_with(tolower)

trait_ref <- pins::pin_get("trait-ref", board = "github")

bt_abund <- bt_data %>%
  filter(abundance_type %in% c("Count", "Density", "MeanCount"))

bt_pres <- bt_data %>%
  filter(abundance_type == "Presence/Absence")
```

```{r}
get_ts_distinct <- function(data, trait_data, ts_id){
  species_data <- data %>% 
    filter(rarefyid == ts_id) %>%
    select(species, abundance, year) %>%
    pivot_wider(names_from = "species", values_from = "abundance", values_fill = 0) %>%
    column_to_rownames("year")

  rel_species <- make_relative(as.matrix(species))
  # traits <- trait_data %>% 
  #   filter(Species %in% unique(species$species)) %>% 
  #   select_if(~length(unique(na.omit(.))) > 1) %>%
  #   column_to_rownames("Species")
  trait_dist <- get_traitMat(colnames(species_data), trait_data)
  #bin_vars <- map(trait_mat, ~all(na.omit(.) %in% 0:1))

  trait_dist <- trait_dist %>%
    mutate_if(is.character, as.factor) %>%
    compute_dist_matrix(., metric = "gower")

  distinctive <- as.data.frame(distinctiveness(pres_matrix = rel_species, dist_matrix = trait_dist)) %>%
    rownames_to_column("year") %>%
    mutate(rarefyid = ts_id)

  return(distinctive)
}

get_ts_distinct(bt_data, trait_ref, "39_407478")

path <- here::here("data", "rarefied_samples")
files <- dir(path, "*.rda") %>%
  paste0(path, "/", .)

plan("multiprocess", workers = 30)
furrr::future_map(files, get_ts_distinct, traits = trait_ref)
```