### 2021-05-13 analysis/visualization.Rmd
Reorganize

---
title: "worldmap"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
library(ggmap)
library(ggplot2)
library(dplyr)
library(ggthemes)
library(mapproj)
library(stringr)
```

```{r}
pins::board_register_github(repo = "karinorman/biodivTS_data", branch = "master")
country_data <- pins::pin_get("country-data-bind", board = "github") %>%
  mutate(country = str_remove(country, "Commonwealth of The "),
         country = str_remove(country, "Kingdom of "),
         country = str_remove(country, "Republic of "),
         country = str_remove(country, "Kingdom of "),
         country = str_remove(country, "Federative Republic of "),
         country = str_remove(country, "Federative "),) %>%
  mutate(country = case_when(
    country == "United States" ~ "USA",
    country == "Russian Federation" ~ "Russia",
    country == "United Great Britain and Northern Ireland" ~ "UK",
    country == "Czechia" ~ "Czech Republic",
    TRUE ~ country
  )) %>% 
  count(country, code)

mapdata <- map_data("world") %>%
  left_join(country_data, by = c("region" = "country"))

mapdata %>%
  ggplot(aes(long, lat)) +
  geom_polygon(mapping = aes(group = group, fill = n)) +
  ggthemes::theme_map() +
  #scale_fill_continuous(color = "#32A251")
  scale_fill_gradient(low = "white", high = "#39737c")

ggsave("study_map.jpeg")
```

```{r}
#get rarefyid centers
included_ids <- pins::pin_get("model-data", board = "github") %>%
  select(rarefyID, taxa, climate, realm) %>%
  distinct() %>% 
  mutate(rarefyID = str_replace_all(rarefyID, pattern = "_bird", ""), 
  rarefyID = str_replace_all(rarefyID, pattern = "_mamm", ""))

study_data <- pins::pin_get("rarefyID-cell-centre", board = "github") %>%
  inner_join(included_ids)

map_data("world") %>%
  ggplot(aes(long, lat, group = group)) +
  geom_polygon(fill = "white", colour = "grey") + 
  coord_map(xlim=c(-180,180), ylim=c(-57, 90)) +
  geom_point(data = study_data, aes(rarefyID_x, rarefyID_y, color = taxa), alpha = 0.5, inherit.aes = FALSE) +
  ggthemes::theme_map()
```