### 2021-08-23 analysis/visualization.Rmd
New color palette, fix overplotting for amphibians.

---
title: "worldmap"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
library(ggmap)
library(ggplot2)
library(dplyr)
library(ggthemes)
library(mapproj)
library(stringr)
library(cowplot)
```

```{r}
pins::board_register_github(repo = "karinorman/biodivTS_data", branch = "master")
# country_data <- pins::pin_get("country-data-bind", board = "github") %>%
#   mutate(country = str_remove(country, "Commonwealth of The "),
#          country = str_remove(country, "Kingdom of "),
#          country = str_remove(country, "Republic of "),
#          country = str_remove(country, "Kingdom of "),
#          country = str_remove(country, "Federative Republic of "),
#          country = str_remove(country, "Federative "),) %>%
#   mutate(country = case_when(
#     country == "United States" ~ "USA",
#     country == "Russian Federation" ~ "Russia",
#     country == "United Great Britain and Northern Ireland" ~ "UK",
#     country == "Czechia" ~ "Czech Republic",
#     TRUE ~ country
#   )) %>% 
#   count(country, code)
# 
# mapdata <- map_data("world") %>%
#   left_join(country_data, by = c("region" = "country"))
# 
# mapdata %>%
#   ggplot(aes(long, lat)) +
#   geom_polygon(mapping = aes(group = group, fill = n)) +
#   ggthemes::theme_map() +
#   #scale_fill_continuous(color = "#32A251")
#   scale_fill_gradient(low = "white", high = "#39737c")

#ggsave("study_map.jpeg")
```

Define color palette
```{r}
my_pal <- c("#d9be3b", "#77976e", "#086788", "#d9be3b", "#99adbf", "#cc8214", "#bfac88")
```

```{r}
#library(forcats)

#get rarefyid centers
included_ids <- pins::pin_get("model-data", board = "github") %>%
  select(rarefyID, taxa, climate, realm, duration) %>%
  distinct() %>% 
  mutate(rarefyID = str_replace_all(rarefyID, pattern = "_bird", ""), 
  rarefyID = str_replace_all(rarefyID, pattern = "_mamm", ""))

study_data <- pins::pin_get("rarefyID-cell-centre", board = "github") %>%
  inner_join(included_ids) #%>%
  #mutate(taxa = as.factor(taxa), taxa = fct_relevel(taxa, "Birds"))

study_map <- map_data("world") %>%
  ggplot(aes(long, lat, group = group)) +
  geom_polygon(fill = "white", colour = "grey") + 
  coord_map(xlim=c(-180,180), ylim=c(-57, 90), projection = "mercator") +
  geom_point(data = study_data,
             aes(rarefyID_x, rarefyID_y, color = taxa), alpha = 0.4, inherit.aes = FALSE) +
  #plot the US amphibian point on top so we can see it
  geom_point(data = study_data %>% 
               #filter(taxa %in% c("Amphibians")), 
               filter(taxa %in% c("Amphibians"), rarefyID_y > 0),
             aes(rarefyID_x, rarefyID_y ), color = "#d9be3b", alpha = 0.8, inherit.aes = FALSE) +
  ggthemes::theme_map() +
  scale_color_manual(values = my_pal) +
  theme(legend.title  = element_blank()) +
  guides(color = guide_legend(override.aes= list(alpha = 0.8)))

ggsave(here::here("figures/study_map.jpeg"))
```

```{r}
duration_hist <- study_data %>% 
  #filter(taxa == "Birds") %>%
  ggplot(aes(duration)) + 
  geom_histogram(aes(fill = taxa)) + 
  theme_classic() + 
  facet_wrap(~taxa, scales = "free") +
  xlab("Timeseries Duration") +
  theme(strip.background = element_blank(), 
        strip.text = element_blank(),
        legend.position = "none", 
        axis.title = element_text(size=8)) +
  scale_fill_manual(values = my_pal)

plot_grid(study_map, duration_hist, labels = "AUTO", nrow = 2, align = 'vh', hjust = -1, axis = "l", scale = c(1,0.9), rel_heights = c(2, 1))

ggsave2(here::here("figures/study_map_hist.jpeg"))
```