### 2021-08-23 analysis/visualization.Rmd
Move timeseries plots from the analysis script, update color palette

---
title: "worldmap"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
library(ggmap)
library(ggplot2)
library(dplyr)
library(ggthemes)
library(mapproj)
library(stringr)
library(cowplot)
library(tidyverse)
```

```{r}
pins::board_register_github(repo = "karinorman/biodivTS_data", branch = "master")
# country_data <- pins::pin_get("country-data-bind", board = "github") %>%
#   mutate(country = str_remove(country, "Commonwealth of The "),
#          country = str_remove(country, "Kingdom of "),
#          country = str_remove(country, "Republic of "),
#          country = str_remove(country, "Kingdom of "),
#          country = str_remove(country, "Federative Republic of "),
#          country = str_remove(country, "Federative "),) %>%
#   mutate(country = case_when(
#     country == "United States" ~ "USA",
#     country == "Russian Federation" ~ "Russia",
#     country == "United Great Britain and Northern Ireland" ~ "UK",
#     country == "Czechia" ~ "Czech Republic",
#     TRUE ~ country
#   )) %>% 
#   count(country, code)
# 
# mapdata <- map_data("world") %>%
#   left_join(country_data, by = c("region" = "country"))
# 
# mapdata %>%
#   ggplot(aes(long, lat)) +
#   geom_polygon(mapping = aes(group = group, fill = n)) +
#   ggthemes::theme_map() +
#   #scale_fill_continuous(color = "#32A251")
#   scale_fill_gradient(low = "white", high = "#39737c")

#ggsave("study_map.jpeg")
```

Define color palette
```{r}
my_pal <- c("#d9be3b", "#77976e", "#086788", "#cc8214", "#bfac88", "#99adbf")
```

```{r}
#library(forcats)

#get rarefyid centers
included_ids <- pins::pin_get("model-data", board = "github") %>%
  select(rarefyID, taxa, climate, realm, duration) %>%
  distinct() %>% 
  mutate(rarefyID = str_replace_all(rarefyID, pattern = "_bird", ""), 
  rarefyID = str_replace_all(rarefyID, pattern = "_mamm", ""))

study_data <- pins::pin_get("rarefyID-cell-centre", board = "github") %>%
  inner_join(included_ids) #%>%
  #mutate(taxa = as.factor(taxa), taxa = fct_relevel(taxa, "Birds"))

study_map <- map_data("world") %>%
  ggplot(aes(long, lat, group = group)) +
  geom_polygon(fill = "white", colour = "grey") + 
  coord_map(xlim=c(-180,180), ylim=c(-57, 90), projection = "mercator") +
  geom_point(data = study_data,
             aes(rarefyID_x, rarefyID_y, color = taxa), alpha = 0.4, inherit.aes = FALSE) +
  #plot the US amphibian point on top so we can see it
  geom_point(data = study_data %>% 
               #filter(taxa %in% c("Amphibians")), 
               filter(taxa %in% c("Amphibians"), rarefyID_y > 0),
             aes(rarefyID_x, rarefyID_y ), color = "#d9be3b", alpha = 0.8, inherit.aes = FALSE) +
  ggthemes::theme_map() +
  scale_color_manual(values = my_pal) +
  theme(legend.title  = element_blank()) +
  guides(color = guide_legend(override.aes= list(alpha = 0.8)))

ggsave(here::here("figures/study_map.jpeg"))
```

```{r}
duration_hist <- study_data %>% 
  #filter(taxa == "Birds") %>%
  ggplot(aes(duration)) + 
  geom_histogram(aes(fill = taxa)) + 
  theme_classic() + 
  facet_wrap(~taxa, scales = "free") +
  xlab("Timeseries Duration") +
  theme(strip.background = element_blank(), 
        strip.text = element_blank(),
        legend.position = "none", 
        axis.title = element_text(size=8)) +
  scale_fill_manual(values = my_pal)

plot_grid(study_map, duration_hist, labels = "AUTO", nrow = 2, align = 'vh', hjust = -1, axis = "l", scale = c(1,0.9), rel_heights = c(2, 1))

ggsave2(here::here("figures/study_map_hist.jpeg"))
```

Metric plots with trend lines
```{r}
metric_model_table <- pins::pin_get("metric-model-table", board = "github")
model_data <- pins::pin_get("model-data", board = "github")

plot_metric <- function(metric_name, ylabel, yval, xlabel, data, model_coef, color_list){

  if(is.na(xlabel)) xlabel <- NULL

  data <- data %>%
    filter(metric == metric_name, !is.na(value), !is.infinite(value))

  model_coef <- model_coef %>%
    filter(metric == metric_name) %>%
    select(term, estimate)

  intercept <- model_coef %>% filter(term == "(Intercept)") %>% pull(estimate)
  slope <- model_coef %>% filter(term == "year_scaled") %>% pull(estimate)

  pred <- data %>%
    select(year_scaled) %>%
    distinct() %>%
    mutate(pred = intercept + year_scaled * slope)

  trend_plot <- data %>%
    left_join(pred, by = "year_scaled") %>%
    #left_join(metadata %>% select(study_id, climate)) %>%
    ggplot(aes(x = year, y = !!sym(yval))) +
    # geom_smooth(aes(color = climate, group = rarefyID), method = "glm", se = FALSE) +
    # geom_point(aes(color = climate, group = rarefyID)) +
    geom_point(aes(group = rarefyID) ,color = "grey", size = 0.25) +
    geom_smooth(aes(color = climate, group = study_id), method = "glm", se = FALSE, size = 0.7) +
    #geom_smooth(color = "black", method = "glm", se = FALSE) +
    geom_line(mapping = aes(y = pred)) +
    theme_classic() +
    theme(axis.title = element_text(size=7), legend.position = "none") +
    ylab(ylabel) +
    xlab(xlabel) +
    labs(color = "Climate") +
    scale_x_continuous(breaks = scales::pretty_breaks(n = 6)) +
    scale_colour_manual(values = color_list, drop = FALSE)

  ggsave(paste0(here::here("figures"), "/",metric_name, "_trend.png"), trend_plot)
  return(trend_plot)
}

model_coef <- metric_model_table %>%
  filter(model_id == "base") %>%
  select(metric, model) %>%
   mutate(coef = map(model, broom.mixed::tidy)) %>% 
  select(-model) %>% 
  unnest(cols = c(coef))

group_colors <- c("Global" = my_pal[1], "Polar/Temperate" = my_pal[5], "Temperate/Tropical" = my_pal[3], "Tropical" = my_pal[4], "Temperate" = my_pal[2], "Overall Mean" = "#000000")

plot_map <- tibble(metric_name = c("S", "Jaccard_base", "SES_FRic", "SES_FDiv", "SES_FEve"), 
           ylabel = c("log(S)", "Jaccard", "Functional Richness SES", "Functional Divergence SES", "Functional Evenness SES"),
           #ylabel = c("log(S)", "Jaccard", "FRic SES", "FDiv SES", "FEve SES"),
           yval = c("logvalue", "value", "value", "value", "value"),
           xlabel = c(NA, NA, NA, "year", "year"))
m_plots <- pmap(plot_map, plot_metric, data = model_data, model_coef = model_coef, color_list = group_colors)

plot_join <- plot_grid(plotlist=m_plots, labels = "AUTO", nrow = 3, align = 'vh',
                       hjust = -1, axis = "l", scale = 0.9)

leg_plot <- model_data %>%
  bind_rows(data.frame(climate = c("Overall Mean", "Overall Mean"), year = c(2019, 2020), value = c(1, 1))) %>%
  ggplot(aes(x = year, y = value)) +
  geom_point(color = "grey", size = 0.50) +
  geom_smooth(aes(color = climate, group = study_id), method = "glm", se = FALSE, size = 0.7) +
  theme_classic() +
  theme(axis.title = element_text(size=7)) +
  scale_colour_manual(values = group_colors, drop = FALSE)

# extract the legend from one of the plots
legend <- get_legend(leg_plot)

# add the legend to the row we made earlier. Give it one-third of 
# the width of one plot (via rel_widths).
#plot_join + draw_grob(legend,  2/3.3, -0.3, .3/3.3, 1, scale = 0.6)
plot_join + draw_grob(legend,  2/3.3, -0.3,.3/3.3, 1, scale = 0.01)

ggsave2(here::here("figures", "3met_long.jpeg"))
```