### 2021-09-16 data-raw/unused_trait_sources/bientraits.Rmd
Rename and rearrange data-raw folder

---
title: "BIEN trait"
author: "Kari Norman"
date: "9/11/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(BIEN)
library(taxadb)

source(here::here("data-raw", "helper_functions.R"))
```

Let's get them traits
```{r}
trait_types <- na.omit(BIEN_trait_list())

BIEN_data <- map_dfr(trait_types$trait_name, BIEN_trait_trait)

bien <- BIEN_data %>%
  select(scientificName = scrubbed_species_binomial, trait_name, trait_value, unit, method, latitude, longitude, elevation_m, obsID = id) %>%
  spread()
```

Get id's for just the unique 
```{r}
bien_ids <- bien %>% 
  select(scientificName) %>%
  distinct() %>%
  mutate(id = get_ids(scientificName))
```

```{r}
usethis::use_data(bien)
```