### 2021-09-16 data-raw/unused_trait_sources/hfp_data.Rmd

---
title: "Human Footprint"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Extract zip files
```{r}
# library(archive)
# hf_arhc <- archive_extract(here::here("data", "HumanFootprintv2.7z.zip"), here::here("data"))
```

Read in raster
```{r}
library(raster)
library(sf)
library(dplyr)
library(geosphere)
library(osmdata)

# met_data <- pins::pin_get("rarefied-metrics", board = "github")
# rarefyIDs <- met_data %>% pull(rarefyID) %>% unique()

hfp_1993 <- raster(here::here("data/Dryadv3/Maps/HFP1993_int.tif")) %>%
  reclassify(., c(100, 128, NA))

hfp_2009 <- raster(here::here("data/Dryadv3/Maps/HFP2009_int.tif")) %>%
  reclassify(., c(100, 128, NA))

#study-level points
# meta <- pins::pin_get("meta", board = "github")
# study_points <- st_as_sf(meta, coords = c("cent_lat", "cent_long"))

#get rarefyid centers
rare_cells <- pins::pin_get("rarefyID-cell-centre", board = "github")
points <- st_as_sf(rare_cells, coords = c("rarefyID_x", "rarefyID_y"), crs = "EPSG:4326") %>%
  st_transform(points, crs = st_crs(hfp_1993)) %>%
  #filter(rarefyID %in% rarefyIDs) %>% 
  dplyr::select(-cell_extent)

# Get 1993 data
ex_1993 <- extract(hfp_1993, points, buffer = 1000, fun = mean, na.rm = TRUE, sp = TRUE) %>%
  st_as_sf(crs = st_crs(hfp_1993))
ex_1993_bigbuff <- ex_1993 %>% filter(is.na(HFP1993_int)) %>% dplyr::select(-HFP1993_int) %>%
  extract(hfp_1993, ., buffer = 45000, fun = mean, na.rm = TRUE, sp = TRUE) %>%
  st_as_sf(crs = st_crs(hfp_1993))

# Get2009 data
ex_2009 <- extract(hfp_2009, points, buffer = 1000, fun = mean, na.rm = TRUE, sp = TRUE) %>%
  st_as_sf(crs = st_crs(hfp_1993))
ex_2009_bigbuff <- ex_2009 %>% filter(is.na(HFP2009_int)) %>% dplyr::select(-HFP2009_int) %>%
  extract(hfp_2009, ., buffer = 45000, fun = mean, na.rm = TRUE, sp = TRUE) %>%
  st_as_sf(crs = st_crs(hfp_1993))

hfp_data <- bind_rows(bind_rows(ex_1993, ex_1993_bigbuff) %>% mutate(hfpYear = 1993), 
          bind_rows(ex_2009, ex_2009_bigbuff) %>% mutate(hfpYear = 2009))

hfp_data <- left_join(bind_rows(ex_1993 %>% filter(!rarefyID %in% ex_1993_bigbuff$rarefyID), ex_1993_bigbuff) %>% 
                        rename(`1993` = HFP1993_int) %>%
                        st_drop_geometry(), 
          bind_rows(ex_2009 %>% filter(!rarefyID %in% ex_2009_bigbuff$rarefyID), ex_2009_bigbuff) %>% 
            rename(`2009` = HFP2009_int) %>%
            st_drop_geometry()) %>%
  pivot_longer(c(`1993`, `2009`), names_to = "year", values_to = "hfp")

pins::pin(hfp_data, board = "github")
```

Attempt at getting distance from shore
```{r}
#Get list of countries
# pins::board_register_github(repo = "karinorman/biodivTS_data", branch = "master")
# country_data <- pins::pin_get("country-data-bind", board = "github") %>%
#   mutate(country = str_remove(country, "Commonwealth of The "),
#          country = str_remove(country, "Kingdom of "),
#          country = str_remove(country, "Republic of "),
#          country = str_remove(country, "Kingdom of "),
#          country = str_remove(country, "Federative Republic of "),
#          country = str_remove(country, "Federative "),) %>%
#   mutate(country = case_when(
#     country == "United States" ~ "USA",
#     country == "Russian Federation" ~ "Russia",
#     country == "United Great Britain and Northern Ireland" ~ "UK",
#     country == "Czechia" ~ "Czech",
#     TRUE ~ country
#   )) 
# 
# map(unique(country_data$country), osm_box)
# 
# get_coasts <- function(country_name) {
#   osm_box <- getbb(place_name = country_name) %>%
#   opq() %>%
#   add_osm_feature("natural", "coastline") %>%
#   osmdata_sf()
# }
# osm_box <- getbb(place_name = "World") %>%
#   opq() %>%
#   add_osm_feature("natural", "coastline") %>%
#   osmdata_sf()
# 
# dist <- geosphere::dist2Line(p = st_as_sf(rare_cells, coords = c("rarefyID_x", "rarefyID_y"), crs = "EPSG:4326") %>% st_coordinates(), 
#                          line = st_coordinates(osm_box$osm_lines)[,1:2])
# 
# #combine initial data with distance to coastline
# df <- cbind( d1 %>% rename(y=lat,x=long),dist) %>%
#   mutate(miles=distance/1609)
# 
# st_as_sf(rare_cells, coords = c("rarefyID_x", "rarefyID_y"), crs = "EPSG:4326")
```

Plot
```{r}
# plot all points 
plot(hfp_1993) 
plot(points, pch = 19, add = TRUE)

plot(hfp_1993) 
plot(ex_1993_bigbuff %>% filter(is.na(HFP1993_int), rarefyID %in% model_data$rarefyID) , pch = 10, add = TRUE)
```