### 2021-09-17 analysis/exporatory_scripts/bbs_timeseries.Rmd
Clean up folders

---
title: "Temporal Analysis"
author: "Kari Norman"
date: "9/19/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#devtools::install_github("karinorman/functional_diversity")

library(tidyverse)
library(dplyr)
library(FD)
library(furrr)
library(here)
library(foreach)
#library(functional.diversity)

piggyback::pb_download(file = c("data/trait.rda", "data/bbs.rda"), repo = "karinorman/functional_diversity", dest = here())

load(here("data", "trait.rda"))
load(here("data", "bbs.rda"))
```

Get FD and TD for each site for each year
```{r}
min_year <- 1969

bbs_wide <- bbs %>%
  select(site_id, scientific, year, abundance) %>%
  spread(scientific, abundance)

bbs_species_mat <- select(bbs_wide, -site_id, -year)
bbs_trait_mat <- get_trait_matrix(colnames(bbs_species_mat))

data_path <- paste0(here('data'), '/FD_bbs_yearly.RData')
  if (file.exists(data_path)){
    print("FD present")
    FD_file <- load(data_path)
  }else{
    print("No FD")
    FD <- as.data.frame(dbFD(bbs_trait_mat, bbs_species_mat, w.abun = TRUE))
    save(FD, file = data_path)
    pb_upload(data_path)
  }

FD <- cbind(bbs_wide[c('site_id', 'year')], FD) #dbFD() preserves row order, so we can just brute force labels back on 
#FD$site_id <- as.factor(FD$site_id)

#double check that labels are right by comparing richnesses calculated below
# richness <- bbs %>%
#   select(site_id, scientific, year) %>%
#   group_by(site_id, year) %>%
#   summarise(n = n())

```

Jarzyna & Jetz 2016 (doi: 10.1111/gcb.13571) calculates a few different metrics to measure temporal change.

Change relative to previous year: (previous year - current year)/current year x 100%
```{r}
yearly_delta <- FD %>%
  select(site_id, year, nbsp, FRic) %>%
  group_by(site_id) %>%
  arrange(year) %>%
  mutate(nbsp_delta = (lag(nbsp) - nbsp)/nbsp,
         fric_delta = (lag(FRic) - FRic)/FRic)
```

Simpson's dissimilarity
Still need to figure out how to deal with holes in the timeseries

Attempt using purrr
```{r}
library(fossil)

get_simpson <- function(site, start_year){
  sp1 <- filter(bbs, site_id == site & year == start_year) %>%
    select(species_id, abundance)

  sp2 <- filter(bbs, site_id == site & year == start_year + 1) %>%
    select(species_id, abundance)

  if(dim(sp2)[1] == 0){
    return(data.frame())
  } else{   
    #need a complete list of species in both sites with zeros for non-occurrences
    comb_sp <- full_join(sp1, sp2, by = "species_id") %>%
      replace_na(list(species_id = 0, abundance.x = 0, abundance.y = 0))

    simp <- simpson(comb_sp$abundance.x, comb_sp$abundance.y)
    return(data.frame(site_id = site, year = start_year + 1, simpson = simp))
  }
}

site_year_comb <- unique(bbs[c('site_id', 'year')]) #get a dataframe of all the site and year combinations that need to be mapped over

bbs_simpson <- map2_dfr(site_year_comb$site_id, site_year_comb$year, get_simpson)
```

For loop approach
```{r}
data_path <- paste0(here('data'), '/simpson_diversity.RData')
if (file.exists(data_path)){
  simpson_file <- load(data_path)
}else{

  site_list <- unique(bbs$site_id)

  #initialize for loops
  simpson_list <- list()
  rownum <- 1

  for (i in 1:length(site_list)) {
    site_data <- filter(bbs, site_id == site_list[i])
    site_years <- unique(site_data$year)

    for (j in 1:length(site_years)) {
      if (j != length(site_years)) { #this will skip the last year which doesn't have a next year to compare to
        sp1 <- filter(site_data, year == site_years[j]) %>%
          select(species_id, abundance)

        sp2 <- filter(site_data, year == site_years[j + 1]) %>%
          select(species_id, abundance)

        comb_sp <- full_join(sp1, sp2, by = "species_id") %>%
          replace_na(list(species_id = 0, abundance.x = 0, abundance.y = 0))

        simp <- simpson(comb_sp$abundance.x, comb_sp$abundance.y)
        simpson_list[[rownum]] <- data.frame(site_id = site_list[i], year = site_years[j+1], simpson = simp)

        rownum <- rownum + 1
      }
    }
  }

  bbs_simpson <- bind_rows(simpson_list)
  save(bbs_simpson, file = data_path)

  pb_upload(data_path)
}
```

Master dataframe of all the metrics
```{r}
metrics <- FD %>%
  select(-starts_with("CWM"), -qual.FRic, -sing.sp) %>%
  left_join(bbs_simpson, by = c("year", "site_id")) %>%
  left_join(yearly_delta, by = c("year", "site_id", "nbsp"))
```

Plotting!
```{r}
sites <- unique(metrics$site_id)
metrics %>%
  filter(site_id %in% sites[1:10]) %>%
  ggplot(aes(year, nbsp, color = as.factor(site_id))) + geom_line() #+
    #theme(legend.position = "none")
```

```{r}
metrics %>%
  group_by(year) %>%
  summarise(ave = mean(nbsp)) %>%
  mutate(ave_5 = RcppRoll::roll_mean(ave, n = 5, fill = NA)) %>%
  ggplot(aes(year, ave_5)) + geom_line()
```