### 2023-01-16 analysis/exporatory_scripts/transformations_outliers.Rmd
First commit

---
title: "Outlier assessment"
output: html_document
date: "2023-01-13"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r setup, include=FALSE}
library(tidyverse)
library(lme4)
library(broom.mixed)
library(lmerTest)
library(emmeans)
```

#### Read in data, get final dataset for models
```{r}
load(here::here("data/model_data.rda"))

model_data <- filter(model_data, taxa != "Amphibians")
```

Let's look at models for FRic

### FRic model

```{r}
fric_data <- model_data %>% filter(metric == "FRic") %>%
  rename(FRic = value) %>%
   mutate(year_scaled = scale(year),
          taxa = as.factor(taxa),
             realm = as.factor(realm),
             climate = as.factor(climate),
          logFRic = log(FRic))
```

let's do the full model for all studies
```{r}
fric_full_fit <- lmer("FRic ~ year_scaled + (year_scaled|study_id/rarefyID)", data = fric_data, REML = FALSE)

logfric_full_fit <- lmer("logFRic ~ year_scaled + (year_scaled|study_id/rarefyID)", data = fric_data, control = lmerControl(optimizer = "bobyqa"))

# remove the outlier
logfric_full_fit_ex <- lmer("logFRic ~ year_scaled + (year_scaled|study_id/rarefyID)", data = fric_data %>% filter(study_id != 515), control = lmerControl(optimizer = "bobyqa"))
```

look at normality
```{r}
# hist(ranef(fric_full_fit)$study_id$`(Intercept)`)
# shapiro.test(ranef(fric_full_fit)$study_id$`(Intercept)`) # test of normality
hist(ranef(fric_full_fit)$study_id$`year_scaled`)
shapiro.test(ranef(fric_full_fit)$study_id$`year_scaled`)

hist(ranef(logfric_full_fit)$study_id$`(Intercept)`)
shapiro.test(ranef(logfric_full_fit)$study_id$`(Intercept)`)
hist(ranef(logfric_full_fit)$study_id$`year_scaled`)
shapiro.test(ranef(logfric_full_fit)$study_id$`year_scaled`)
```
```{r}
test_normality <- function(model_fit, metric_name){
  intercepts <- ranef(model_fit)$study_id$`(Intercept)`
  slopes <- ranef(model_fit)$study_id$`year_scaled`

  hist(intercepts, main = paste("intercepts for", metric_name))
  hist(slopes, main = paste("slopes for", metric_name))

  bind_rows(tidy(shapiro.test(intercepts)) %>% mutate(type = "intercept"),
            tidy(shapiro.test(slopes)) %>% mutate(type = "slope")) %>%
    mutate(metric = metric_name)
}

test_normality(logfric_full_fit, "log(FRic)")
```

# Look at normality for base models we've already fit

```{r}
# model_table comes from 05_model_metrics.Rmd

model_df <- model_table %>%
  filter(model_id == "base") %>%
  select(metric, model, model_id) %>%
  group_by(metric, model_id) %>% 
  slice_head()

mod_normality <- map2_dfr(model_df$model, model_df$metric, test_normality)
```

# Let's make almost everything log, not SES metrics because they're already normalized
```{r}
metric_names <- unique(model_data$metric)

log_mods <- tibble(metric_name = metric_names[!metric_names %in% c("SES_FRic", "SES_FDiv", "SES_FEve")], 
                         formula = "logvalue ~ year_scaled + (year_scaled|study_id/rarefyID)",
                   model_id = "base")

log_mods_fit <- pmap_dfr(log_mods, get_model, data = model_data)

log_model_df <- log_mods_fit %>%
  filter(model_id == "base") %>%
  select(metric, model, model_id) %>%
  group_by(metric, model_id) %>% 
  slice_head()

logmod_normality <- map2_dfr(log_model_df$model, log_model_df$metric, test_normality)
```

# The logvalue in model_data is actually log(value + 1), which works for the CWM but not the summary metrics
```{r}
log_metrics <- c("FRic", "S", "FDiv", "FEve")
log_mods <- tibble(metric_name = log_metrics, 
                         formula = "log(value) ~ year_scaled + (year_scaled|study_id/rarefyID)",
                   model_id = "base")

logvalue_mods <- tibble(metric_name = metric_names[!metric_names %in% append(c("SES_FRic", "SES_FDiv", "SES_FEve"), log_metrics)], 
                         formula = "logvalue ~ year_scaled + (year_scaled|study_id/rarefyID)",
                   model_id = "base")

trans_mods <- bind_rows(log_mods, logvalue_mods)

trans_mods_fit <- pmap_dfr(trans_mods, get_model, data = model_data) %>%
  mutate(REML = TRUE, bobyqa = FALSE)

# have to make sure they all converge #

#get the models that had convergence warnings
bad_models <- trans_mods_fit %>%
  select(metric, model_id, warning) %>% 
  filter(!is.na(warning)) %>% 
  distinct() %>%
  left_join(trans_mods, by = c("metric" = "metric_name", "model_id"))

#Turn off REML for these models, and they fit! 
bad_mod_results <- pmap_dfr(bad_models %>% select(-warning), get_model, data = model_data, REML = FALSE) %>%
  mutate(REML = FALSE, bobyqa = FALSE)

#models that still won't converge
bad_models2 <- bad_mod_results %>%
  select(metric, model_id, warning) %>% 
  filter(!is.na(warning)) %>% 
  distinct() %>%
  left_join(trans_mods, by = c("metric" = "metric_name", "model_id"))

#Fit with a different algorithm
bad_mod_results2 <- pmap_dfr(bad_models2 %>% select(-warning), get_model, data = model_data, control = lmerControl(optimizer = "bobyqa")) %>%
  mutate(REML = TRUE, bobyqa = TRUE)

transmodel_table <- bind_rows(trans_mods_fit %>% filter(is.na(warning)), 
          bad_mod_results %>% filter(is.na(warning)),
          bad_mod_results2 %>% filter(is.na(warning)))

# let's look at normality of the random effects
trans_model_df <- transmodel_table %>%
  select(metric, model, model_id, formula, REML, bobyqa) %>%
  group_by(metric, model_id) %>% 
  slice_head() %>%
  ungroup()

transmod_normality <- map2_dfr(trans_model_df$model, trans_model_df$metric, test_normality)
```

# Let's figure out if we have outlier issues, after visually inspecting the random effects

```{r}
s_mod <- trans_model_df %>% filter(metric == "S") %>% pull(model) %>% .[[1]]

intercepts <- ranef(s_mod)$study_id$`(Intercept)`
slopes <- ranef(s_mod)$study_id$`year_scaled`

which(slopes < -1) # It's row 22
ranef(s_mod)$study_id[22,] # It's study 377

s_fit <- lmer("log(value) ~ year_scaled + (year_scaled|study_id/rarefyID)", data = model_data %>% filter(metric == "S", study_id != 377), control = lmerControl(optimizer = "bobyqa"))
```

Let's make it a function
```{r}
check_outliers <- function(start_mod, metric_name, outlier_criterion, lmer_formula, REML = TRUE, bobyqa = FALSE){

  intercepts <- ranef(start_mod)$study_id$`(Intercept)`
  slopes <- ranef(start_mod)$study_id$`year_scaled`

  rows <- which(eval(rlang::parse_expr(outlier_criterion))) # It's row 22
  print(rows)
  if(length(rows) > 1) warning(paste("There is more than one study outside the outlier cutoff for", metric_name))

  study <- as.integer(rownames(ranef(start_mod)$study_id[rows,]))

  if (isTRUE(REML)){
    if (isTRUE(bobyqa)) {
        fit <- lmer(lmer_formula, 
              data = model_data %>% filter(metric == metric_name, study_id != study), 
              control = lmerControl(optimizer = "bobyqa"))
    } else {
      fit <- lmer(lmer_formula, 
              data = model_data %>% filter(metric == metric_name, study_id != study))
    }
  } else{
    if (isTRUE(bobyqa)) {
        fit <- lmer(lmer_formula, 
              data = model_data %>% filter(metric == metric_name, study_id != study), 
              REML = FALSE,
              control = lmerControl(optimizer = "bobyqa"))
    } else {
      fit <- lmer(lmer_formula, 
              data = model_data %>% filter(metric == metric_name, study_id != study),
              REML = FALSE)
    }
  }

  tibble(metric = metric_name,
         old_estimate = tidy(start_mod) %>% filter(term == "year_scaled") %>% pull(estimate), 
         old_p = tidy(start_mod) %>% filter(term == "year_scaled") %>% pull(p.value), 
         new_estimate = tidy(fit) %>% filter(term == "year_scaled") %>% pull(estimate), 
         new_p = tidy(fit) %>% filter(term == "year_scaled") %>% pull(p.value))

}

#check_outliers("S", "slopes < -1", "log(value) ~ year_scaled + (year_scaled|study_id/rarefyID)", REML = TRUE, bobyqa = TRUE)

# create dataframe for outlier criterion based on visualize of random effects
criterion <- tibble(metric_name = c("CWM_bodymass_value", "CWM_diet_inv", "CWM_diet_scav", "CWM_diet_vect", 
                                     "CWM_diet_vend", "CWM_diet_vunk", "CWM_forstrat_aerial", "CWM_forstrat_midhigh",
                                     "CWM_forstrat_understory", "Jaccard_base", "S"),
                    outlier_criterion = c("slopes > 0.4", "slopes < -0.5", "slopes > 0.6", "slopes > 0.4", 
                                          "slopes > 0.5", "slopes < -0.5", "slopes > 0.04", "slopes < -0.05", 
                                          "slopes < -0.4", "slopes > 0.02", "slopes < -1")
                    )

# get metric info for how models were fit
outlier_mods <- trans_model_df %>% 
  select(start_mod = model, metric_name = metric, lmer_formula = formula, REML, bobyqa) %>%
  right_join(criterion) %>%
  select(start_mod, metric_name, outlier_criterion, lmer_formula, REML, bobyqa)

outlier_comp <- pmap_dfr(outlier_mods, check_outliers)

```