### 2023-02-14 analysis/exporatory_scripts/compare_transformations.Rmd

---
title: "Outlier assessment"
output: html_document
date: "2023-01-13"
---

```{r setup, include=FALSE}
library(tidyverse)
library(lme4)
library(broom.mixed)
library(lmerTest)
library(emmeans)
```

#### Read in data, get final dataset for models
```{r}
load(here::here("data/model_data.rda"))

model_data <- filter(model_data, taxa != "Amphibians")
```

Create dataframe of mixed effects models to fit
```{r}
# metric_names <- unique(model_data$metric)
# raw_metrics <- c("SES_FRic", "SES_FDiv", "SES_FEve")
# log_metrics <- c("FRic", "S", "FDiv", "FEve")
# logone_metrics <- metric_names[!metric_names %in% c(raw_metrics, log_metrics)]
# 
# raw_value_mods <- tibble(metric_name = raw_metrics, 
#                          base = "value ~ year_scaled + (year_scaled|study_id/rarefyID)",
#                          taxa = "value ~ year_scaled * taxa + (year_scaled|study_id/rarefyID)",
#                          realm = "value ~ year_scaled * realm + (year_scaled|study_id/rarefyID)",
#                          climate = "value ~ year_scaled * climate + (year_scaled|study_id/rarefyID)") %>%
#   pivot_longer(cols = c(base, taxa, realm, climate), names_to = "model_id", values_to = "formula") %>%
#   #don't fit the CWM_litter_size_min_n and climate model, not enough data
#   filter(metric_name != "CWM_litter_size_min_n" | model_id != "climate")
# 
# log_mods <- tibble(metric_name = log_metrics, 
#                          base = "log(value) ~ year_scaled + (year_scaled|study_id/rarefyID)",
#                          taxa = "log(value) ~ year_scaled * taxa + (year_scaled|study_id/rarefyID)",
#                          realm = "log(value) ~ year_scaled * realm + (year_scaled|study_id/rarefyID)",
#                          climate = "log(value) ~ year_scaled * climate + (year_scaled|study_id/rarefyID)") %>%
#   pivot_longer(cols = c(base, taxa, realm, climate), names_to = "model_id", values_to = "formula")
# 
# logone_mods <- tibble(metric_name = logone_metrics, 
#                          base = "logvalue ~ year_scaled + (year_scaled|study_id/rarefyID)",
#                          taxa = "logvalue ~ year_scaled * taxa + (year_scaled|study_id/rarefyID)",
#                          realm = "logvalue ~ year_scaled * realm + (year_scaled|study_id/rarefyID)",
#                          climate = "logvalue ~ year_scaled * climate + (year_scaled|study_id/rarefyID)") %>%
#   pivot_longer(cols = c(base, taxa, realm, climate), names_to = "model_id", values_to = "formula")
# 
# mods <- bind_rows(raw_value_mods, log_mods, logone_mods) %>% 
#   select(metric_name, formula, model_id)
```

Fit models
```{r}
# #exclude amphibians
# model_data <- model_data %>% filter(taxa != "Amphibians")
# 
# #Fit models with base settings
# mod_results <- pmap_dfr(mods, get_model, data = model_data) %>%
#   mutate(REML = TRUE, bobyqa = FALSE)
# 
# #get the models that had convergence warnings
# bad_models <- mod_results %>%
#   select(metric, model_id, warning) %>% 
#   filter(!is.na(warning)) %>% 
#   distinct() %>%
#   left_join(mods, by = c("metric" = "metric_name", "model_id")) 
# 
# #Turn off REML for these models, and they fit! 
# bad_mod_results <- pmap_dfr(bad_models %>% select(-warning), get_model, data = model_data, REML = FALSE) %>%
#   mutate(REML = FALSE, bobyqa = FALSE)
# 
# #models that still won't converge
# bad_models2 <- bad_mod_results %>%
#   select(metric, model_id, warning) %>% 
#   filter(!is.na(warning)) %>% 
#   distinct() %>%
#   left_join(mods, by = c("metric" = "metric_name", "model_id")) 
# 
# #Fit with a different algorithm
# bad_mod_results2 <- pmap_dfr(bad_models2 %>% select(-warning), get_model, data = model_data, bobyqa = TRUE) %>%
#   mutate(REML = TRUE, bobyqa = TRUE)
# 
# model_table <- bind_rows(mod_results %>% filter(is.na(warning)), 
#           bad_mod_results %>% filter(is.na(warning)),
#           bad_mod_results2 %>% filter(is.na(warning))) %>%
#   mutate(significant = case_when(
#     p.value > 0.05 | lower.CL < 0 & upper.CL > 0 ~ FALSE,
#     !is.na(p.value) | !is.na(lower.CL) ~ TRUE,
#     TRUE ~ NA
#   )) 
```

Need to also fit the non-log models
```{r}
nonlog_mods <- tibble(metric_name = metric_names[metric_names != "S"], 
                         base = "value ~ year_scaled + (year_scaled|study_id/rarefyID)",
                         taxa = "value ~ year_scaled * taxa + (year_scaled|study_id/rarefyID)",
                         realm = "value ~ year_scaled * realm + (year_scaled|study_id/rarefyID)",
                         climate = "value ~ year_scaled * climate + (year_scaled|study_id/rarefyID)") %>%
  pivot_longer(cols = c(base, taxa, realm, climate), names_to = "model_id", values_to = "formula") %>%
  #don't fit the CWM_litter_size_min_n and climate model, not enough data
  filter(metric_name != "CWM_litter_size_min_n" | model_id != "climate") %>% 
  select(metric_name, formula, model_id)
```

Fit models
```{r}
#Fit models with base settings
nonlog_mod_results <- pmap_dfr(nonlog_mods, get_model, data = model_data)

#get the models that had convergence warnings
nonlog_bad_models <- nonlog_mod_results %>%
  select(metric, model_id, warning) %>% 
  filter(!is.na(warning)) %>% 
  distinct() %>%
  left_join(mods, by = c("metric" = "metric_name", "model_id"))

#Turn off REML for these models, and they fit! 
nonlog_bad_mod_results <- pmap_dfr(nonlog_bad_models %>% select(-warning), get_model, data = model_data, REML = FALSE)

#models that still won't converge
nonlog_bad_models2 <- nonlog_bad_mod_results %>%
  select(metric, model_id, warning) %>% 
  filter(!is.na(warning)) %>% 
  distinct() %>%
  left_join(mods, by = c("metric" = "metric_name", "model_id"))

#Fit with a different algorithm
nonlog_bad_mod_results2 <- pmap_dfr(nonlog_bad_models2 %>% select(-warning), get_model, data = model_data, control = lmerControl(optimizer = "bobyqa"))

nonlog_model_table <- bind_rows(nonlog_mod_results %>% filter(is.na(warning)), 
          nonlog_bad_mod_results %>% filter(is.na(warning)),
          nonlog_bad_mod_results2 %>% filter(is.na(warning))) %>%
  mutate(significant = case_when(
    p.value > 0.05 | lower.CL < 0 & upper.CL > 0 ~ FALSE,
    !is.na(p.value) | !is.na(lower.CL) ~ TRUE,
    TRUE ~ NA
  ))
```

Function to compare pre and post transformation normality
```{r}
library(DHARMa)

comp_normality <- function(metric, model, log_model){
  model_sim <- simulateResiduals(fittedModel = model, plot = F)
  log_model_sim <- simulateResiduals(fittedModel = log_model, plot = F)

  plot(model_sim, title = paste(metric, "untransformed"))
  plot(log_model_sim, title = paste(metric, "log transformed"))
}

model_comp_df <- nonlog_model_table %>%
  filter(model_id == "base") %>%
  select(metric, model) %>% 
  group_by(metric) %>%
  slice_head(n=1) %>%
  left_join(model_table %>% 
              filter(model_id == "base") %>%
              select(metric, log_model = model) %>% 
              group_by(metric) %>%
              slice_head(n=1))

pmap(model_comp_df, comp_normality)
```
Based on visual comparison, the following metrics are better *not* transformed: diet_inv, forstrat_ground, forstrat_watbelowsurf, FEve.