### 2023-02-14 analysis/exporatory_scripts/slope_est_comparison.Rmd

---
title: "Untitled"
output: html_document
date: "2023-02-08"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Get study level slope estimates using BLUP's
```{r}
get_study_slopes <- function(metric, model, model_id){

  filter_name <- ifelse(model_id == "base", "study_id", model_id)

  #get the statistics for the year variable
  year_scaled_coef <- tidy(model) %>%
    filter(term == "year_scaled")

  #get the year estimate
  year_scaled_est <- pull(year_scaled_coef, estimate)

  #get year variation
  year_scaled_var <- year_scaled_coef %>% 
    mutate(var = std.error*std.error) %>%
    pull(var)

  #get study level conditional estimates and add to over all estimates
  study_ests <- broom.mixed::tidy(model, effects="ran_vals") %>%
    filter(group == filter_name, term == "year_scaled") %>% 
    select(-c(effect, group, term)) %>%
    rename(cond.std.error = std.error, cond.estimate = estimate) %>%
    mutate(estimate = cond.estimate + year_scaled_est,
      cond.var = cond.std.error*cond.std.error, 
           var = cond.var + year_scaled_var,
           std.error = sqrt(var),
           upr.ci = estimate + (1.96*std.error),
           lwr.ci = estimate - (1.96*std.error),
           sig = case_when(
             lwr.ci < 0 & upr.ci > 0 ~ FALSE,
             TRUE ~ TRUE
           ),
      metric = metric,
      model_id = model_id)
}

model_df <- model_table %>%
  filter(model_id == "base") %>%
  select(metric, model, model_id) %>%
  group_by(metric, model_id) %>% 
  slice_head()

#get study level slopes for base, and levels for categorical random effect in non-base models
re_slopes <- pmap_dfr(model_df, get_study_slopes)

#get only study-level slopes for the base models
study_slopes <- re_slopes %>% 
  filter(model_id == "base") %>%
  rename(study_id = level)

usethis::use_data(study_slopes)
```

Get study level slopes using bootstrap resampling
```{r}
extract_study_slopes <- function(model){

  #get the statistics for the year variable
  year_scaled_est <- tidy(model) %>%
    filter(term == "year_scaled") %>%
    pull(estimate)

  estimate <- broom.mixed::tidy(model, effects="ran_vals") %>%
    filter(group == "study_id", term == "year_scaled") %>% 
    select(-c(effect, group, term)) %>%
    rename(cond.estimate = estimate) %>%
    mutate(estimate = cond.estimate + year_scaled_est) %>%
    select(level, estimate) %>%
    deframe()

  return(estimate)
}

bootstrap_study_slopes <- function(metric, model, model_id, nsim){

  print(c(metric, model_id))

  fric_vars <- bootMer(model, FUN = extract_study_slopes, nsim = nsim, use.u = TRUE)

  estimates <- enframe(fric_vars$t0, name = "study_id", value = "estimate")

  as_tibble(fric_vars$t) %>% 
    pivot_longer(everything(), names_to = "study_id", values_to = "values") %>%
    group_by(study_id) %>%
    summarize(n = n(), std.error = sd(values)/sqrt(n)) %>%
    left_join(estimates) %>%
    select(-n) %>%
    mutate(metric = metric, model_id = model_id)
}

#future::plan("multisession", workers = 30)

bootstrap_slopes <- pmap_dfr(model_df %>% filter(model_id == "base") %>% mutate(nsim = 100), bootstrap_study_slopes) %>%
  # furrr::future_pmap_dfr(model_df %>% filter(model_id == "base") %>% mutate(nsim = 100), bootstrap_study_slopes, 
  #                                          .options = furrr::furrr_options(seed = TRUE)) %>%
  mutate(upr.ci = estimate + (2*std.error),
         lwr.ci = estimate - (2*std.error),
         sig = case_when(
             lwr.ci < 0 & upr.ci > 0 ~ FALSE,
             TRUE ~ TRUE
           ))
```

```{r}
slope_comp <- study_slopes %>% 
  select(study_id, metric, blup_estimate = estimate, blup_stde = std.error, sig_bulp = sig) %>%
  left_join(bootstrap_slopes %>% 
              select(study_id, metric, boot_estimate = estimate, boot_stde = std.error, sig_boot = sig)) %>%
  left_join(indv_mod_results %>%
              filter(term == "year_scaled") %>%
            mutate(sig_indv = if_else(p.value < 0.05, TRUE, FALSE)) %>%
              select(study_id, metric, indv_estimate = estimate, indv_stde = std.error, sig_indv))
```

```{r}
load(here::here("data/indv_mod_table.rda"))

slopes_long <- slope_comp %>% 
    select(study_id, metric, blup_estimate, blup_stde, boot_estimate, boot_stde, indv_estimate, indv_stde) %>%
  pivot_longer(cols = c(blup_estimate, blup_stde, boot_estimate, boot_stde, indv_estimate, indv_stde),
  names_to = c("type", ".value"),
  names_pattern = "(.+)_(.+)"
  )

slopes_long %>% 
  #filter(metric %in% c("S", "Jaccard_base", "SES_FRic", "SES_FDiv", "SES_FEve")) %>%
  filter(metric == "Jaccard_base") %>%
  filter(type != "blup") %>%
  ggplot(aes(x = study_id, y = estimate)) +
  geom_point(aes(colour = type), size = 3.5) +
  geom_errorbar(aes(ymin = estimate-(2*stde), ymax = estimate + (2*stde))) + 
    facet_wrap(vars(metric))

```