### 2023-06-05 06_study_change.Rmd
Get the table working for knit at least!

---
title: "Study Level trends"
output: pdf_document
header-includes:
  - \usepackage{booktabs}
date: "2023-06-05"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
```

```{r}
load(here::here("data/indv_mod_table.rda"))
load(here::here("data/model_data.rda"))
```

# Types of Biodiversity Change

```{r}
wide_table <- indv_mod_table %>% 
  filter(p.value < 0.05, term == "year_scaled", model_id == "base") %>% 
  dplyr::select(study_id, metric, estimate) %>% 
  pivot_wider(names_from = "metric", values_from = "estimate") %>% 
  dplyr::select(., -starts_with("CWM_")) %>% 
  dplyr::select(study_id, S, Jaccard_base, everything(), -Jaccard_next) %>%
  left_join(model_data %>% 
              dplyr::select(study_id, duration) %>%
              distinct() %>%
              group_by(study_id) %>% 
              mutate(duration = max(duration)) %>% distinct() %>% 
              mutate(study_id = as.character(study_id))) 
```
# Add decriptive columns to table 
```{r}
fun_change_table <- wide_table %>%
  mutate(rich_change = ifelse(!is.na(S), "Richness Change", "No Richness Change"),
         turnover = ifelse(!is.na(Jaccard_base), "Turnover", "No Turnover"),
         fun_change = case_when(
           if_all(c(FRic, SES_FRic, FDiv, SES_FDiv, FEve, SES_FEve), ~is.na(.)) ~ "No Functional Change",
           .default = "Functional Change"
         ))

labels = fun_change_table %>%
  dplyr::select(rich_change, turnover, fun_change) %>%
  distinct() %>%
  mutate(group = LETTERS[1:8])

fun_change_table <- fun_change_table %>%
  left_join(labels)

fun_change_table %>% 
  count(rich_change, turnover, fun_change, group) 
```

Let's split up the corrected and uncorrected functional metrics
```{r}
ses_wide_table <-  wide_table %>%
  dplyr::select(-FRic, -FDiv, -FEve) %>%
  mutate(rich_change = ifelse(!is.na(S), TRUE, FALSE),
         turnover = ifelse(!is.na(Jaccard_base), TRUE, FALSE),
         fun_change = case_when(
           if_all(c(SES_FRic, SES_FDiv, SES_FEve), ~is.na(.)) ~ FALSE,
           .default = TRUE
         ))

ses_wide_table %>% 
  count(rich_change, turnover, fun_change) 
```
```{r}
uncor_wide_table <-  wide_table %>%
  dplyr::select(-SES_FRic, -SES_FDiv, -SES_FEve) %>%
  mutate(rich_change = ifelse(!is.na(S), TRUE, FALSE),
         turnover = ifelse(!is.na(Jaccard_base), TRUE, FALSE),
         fun_change = case_when(
           if_all(c(FRic, FDiv, FEve), ~is.na(.)) ~ FALSE,
           .default = TRUE
         ))

uncor_wide_table %>% 
  count(rich_change, turnover, fun_change) 
```

# Look at distribution of durations
```{r}
duration_table <- fun_change_table %>%
  dplyr::select(study_id, rich_change, turnover, fun_change, group) %>%
  left_join(model_data %>% 
              dplyr::select(study_id, rarefyID, duration) %>% 
              distinct())

```

```{r}
plot_table <- duration_table %>%
  filter(group %in% c("A", "B", "C", "D"))

group_types <- as_labeller(
  c("A" = "No richness change or turnover, functional change", "B" = "No turnover, richness and functional change", "C" = "No change", "D" = "Turnover, richness and functional Change")
)

my_pal <- c("#d9be3b", "#77976e", "#086788", "#cc8214", "#bfac88", "#99adbf")
group_colors <- c("A" = "#d9be3b", "B" = "#77976e", "C" = "#086788", "D" = "#cc8214")

ggplot(plot_table, aes(x = duration, fill = group)) +
  geom_histogram(data = dplyr::select(plot_table, -group), fill = "grey", alpha = .5) +
  geom_histogram(colour = "black") +
  facet_wrap(~ group, labeller = group_types) +
  guides(fill = FALSE) +  # to remove the legend
  scale_fill_manual(values = group_colors) +
  theme_classic()      
```
Let's make a pretty table now
```{r, echo=FALSE}
library(kableExtra)

fun_change_table %>% 
  dplyr::select(-group) %>%
  count(rich_change, turnover, fun_change) %>%
  pivot_wider(names_from = rich_change, values_from = n) %>%
  select(fun_change, everything()) %>%
  arrange(fun_change) %>%
  mutate(`No Richness Change` = cell_spec(`No Richness Change`,
                                           color = case_when(
                                             `No Richness Change` == 12 ~ "#d9be3b",
                                             `No Richness Change` == 11 ~ "#086788",
                                             .default = "black"))) %>%
  mutate(`Richness Change` = cell_spec(`Richness Change`,
                                           color = case_when(
                                             `Richness Change` == 6 ~ "#77976e",
                                             `Richness Change` == 7 ~ "#cc8214",
                                             .default = "black"))
         ) %>%
  kable(format = "latex", col.names = c("", "", "No Richness Change", "Richness Change"), 
        booktabs = TRUE, align = "c", escape = FALSE) %>%
  kable_styling("striped") %>%
  row_spec(0, bold = TRUE) %>%
  column_spec(1:2, bold = TRUE) %>%
  collapse_rows(columns = 1, valign = "middle") %>%
  kable_as_image("figures/change_table", file_format = "jpeg", latex_header_includes = c("\\\\usepackage{booktabs}"))
```