### 2024-05-23 paper/GEB Submission/manuscript_GEB.Rmd
Additional intro paragraph, other minor edits from Ben.

---
title: No general trend for multiple dimensions of functional diversity in bird and mammal communities
abstract: |
  Aim: Despite unprecedented environmental change due to anthropogenic pressure, recent work has found increasing dissimilarity due to turnover but no overall trend in species diversity through time at the local scale. Functional diversity provides a potentially powerful alternative approach for understanding community composition by linking shifts in species identity to the characteristics that underpin ecosystem processes. Here we present the first multi-taxa, multi-system analysis of functional change through time.

  Location: Global, with a North American focus

  Time period: 1923-2014

  Major taxa studied: Mammals, Birds

  Methods: We paired thousands of bird and mammal assemblage time series from the BioTIME database with existing trait data representative of species' functional roles to reconstruct time series of functional diversity and composition metrics. Using generalized linear mixed models, we estimated general trends in those metrics and trends for individual studies.

  Results: We found no overall temporal trends in functional diversity metrics. Study characteristics such as taxa, realm, biome, or protection status did not distinguish between types of change exhibited by communities. We found evidence of an increasing temporal trend in fish consumption across all communities with , but no evidence to support multiple predictions for specific traits, including decreasing body size, dietary shifts, or changes in foraging strata.

  Main Conclusions: General temporal trends indicate that on the aggregate one type of functional shift is not more prevalant than the other across many taxa, biomes, and realms. At the study-level, the majority of studies showed no species or functional trends, with the remainder of the studies falling into broad categories of functional shifts independent of species richness, functional redundancy loss, and increasing species richness.

journal: "Ecology Letters"
date: "`r Sys.Date()`"
bibliography: refs.bib
csl: pnas.csl
layout: 3p
header-includes:
   - \usepackage{lineno}
   - \linenumbers
output: 
  rticles::arxiv_article:
    keep_tex: true
#  word_document: default
editor_options: 
  markdown: 
    wrap: sentence
keywords:
  - biodiversity change
  - functional traits
  - global change
  - time series

nocite: |
  @bakker1990, @berezovikovn.n.2004, @enemar2004, @gido2019, @hogstad1993, @holmes1986, @jahncke2006, @krivenkov.g.1991, @melnikovy.i.2000, @monitoringsite1000projectbiodiversitycenterministryofenvironmentofjapan2013, @monitoringsite1000projectbiodiversitycenterministryofenvironmentofjapan2014, @monitori2014, @pardieck2020, @preston1960, @svensson2006, @svensson2010, @thorn2016changes, @thorn2016response, @usfs, @vermontcenterforecostudies2015, @vickery1984, @waide2017, @williamson1983, @woehlerericj., @zakharovv.d.1998, @meditss2011, @animald, @baltics, @piropno, @rossaferes1997, @scott, @silva, @ernest2009, @friggens2008, @jalilova.b.2014, @kartzinel2014, @kaufman2019, @krefting1974, @lightfoot, @malyshevy.s.2011, @meyer2008, @nedosekinv.y.2015, @prins1990, @rocha2017, @sheftelb.i., @stapp2014, @monitori2014, @bahamas, @popacet
---

```{r message=FALSE, include = FALSE}
library(kableExtra)
library(magrittr)
library(tidyverse)
#library(printr)
library(rticles)
library(knitr)

knitr::opts_chunk$set(cache=FALSE, message=FALSE, warning=FALSE, out.width = "\\textwidth")
options(knitr.kable.NA = '')

# no sci notation integers pleases
options(scipen=999)

colorize <- function(x, color) {
  if (knitr::is_latex_output()) {
    sprintf("\\textcolor{%s}{%s}", color, x)
  } else if (knitr::is_latex_output()) {
    sprintf("<span style='color: %s;'>%s</span>", color,
      x)
  } else x
}
```

# Introduction

Ecological communities are experiencing unprecedented changes as a result of anthropogenic pressures such as climate change, land use change, and invasive species.
Impacts of these pressures are well documented at a global scale by an accelerating global extinction rate [@barnosky2011] and fundamental changes in some of the most well-studied systems [e.g. coral bleaching, @sully2019].
At the local scale however, species diversity tells a different story.
Recent syntheses of local trends in biodiversity over time have found no net change in local species diversity despite ongoing turnover [@brown2001; @dornelas2014; @vellend2013; @vellend2017] and evidence of significant shifts in community composition underlying consistent species richness [@brose2016; @gotelli2017; @li2020].
While communities are clearly changing, our most common species-based approaches do not fully capture the nature of that change.

`r colorize("Using general trends derived from limited data as a diagnostic for the state of biodiversity is a topic of ongoing debate.", "blue")` Global analyses have been heavily criticized for geographic biases, lack of data in the most heavily impacted areas, and exclusion of individual studies' ecological context [@cardinale2014; @cardinale2018; @gonzalez2016].
Many of these criticisms reflect limitations of ecological data on the whole, `r colorize("leading to a call for additional data not only to fill geographic and temporal gaps, but to fill in key characteristics of communities", "blue")` [@dornelas2023; @primack2018].

Functional diversity offers a potentially powerful addition to species-based approaches for detecting and describing community change by providing a mechanistic link between species' response to environmental change (*response traits*) and the processes they perform (*effect traits*) [@lavorel2002; @mcgill2006; @suding2008].
By describing the functional trait space, functional diversity metrics capture the disproportionate impact of losses or gains of functionally unique species.
Functional diversity metrics `r colorize("may therefore illuminate joint responses from functionally similar species or communities undetectible by looking at species identity alone.", "blue")`

Broad-scale trends in functional diversity are the aggregation of local-scale responses to disturbance and persistent pressures like environmental change or the introduction of invasive species.
On acute timescales, impacts on communities can range from reduction in abundance across the trait space to complete gains or losses in some aspect of the trait space [@boersma2016; @larson2021].
As communities recover or adapt to persistent pressures, changes in the functional space are determined by community assembly processes and limited by the species available in the regional species pool [@green2022].
Lost species may be replaced by functionally similar species when available, or novel species may colonize niches newly created by changing environmental conditions.
Critically, similar shifts in the functional space may be the result of both positive and negative conservation scenarios.
For example, expansion of the functional trait space may be evidence of recovery from disturbance or the introduction of a functionally unique invasive species.

`r colorize("The expectation for functional diversity change across communities is not obvious from past work, and may or may not follow species richness trends", "blue")` [@dornelas2023].
While `r colorize("loss of functional diversity", "blue")` is frequently cited as one of the most pressing concerns of the anthropocene [@cardinale2012; @dirzo2014; @young2016], `r colorize("functional diversity may be maintained even when species are lost from a community", "blue")` [@daz2001; @petchey2009].
Forecasts of functional loss range from negligible [@gallagher2013] to dire [@petchey2002; @pimiento2020].
And while some observed trends show significant functional loss [@flynn2009] others document no loss even in some of the most heavily impacted communities [@edwards2013; @larsen2018; @matuoka2020].
On paleoecological time scales, `r colorize("the functional space", "blue")` shows mixed responses to environmental change and extinction events [@jackson2015; @dornelas2018], with significant impacts of species extinctions on functional diversity in some taxa and not others [@pimiento2017].
For some time periods, functional diversity appears to be maintained for substantial portions of geological time [@hedberg2022].
`r colorize("Contemporary, broad-scale examinations of functional change are limited to only a few taxa-focused studies, but show for example functional richness increases for both North American birds", "blue")` [@barnagaud2017; @jarzyna2016] `r colorize("and ray-finned fishes, sharks, and rays", "blue")` [@trindade-santos2020].

`r colorize("We have stronger expectations for changes in the prevalence of some individual traits. For example, animal body size is expected to decrease as a result of climate change, a phenomena that has been documented in multiple taxa empirically and experimentally", "blue")` [@caruso2014; @forster2012; @huss2019; @sheridan2011; @tseng2018].
`r colorize("There is some evidence that this holds true for birds ", "blue")`[@vanbuskirk2010]`r colorize(", but the picture is likely more complicated for mammals, where urbanization may actually lead to larger body sizes due to novel food sources", "blue")` [@hantak2021] `r colorize(", even as megafaunal loss leads to on average smaller community body size. For dietary traits, recent work documenting insect declines", "blue")` [@wagner2021] `r colorize("points to potentially significant negative impacts on insectivorous animals", "blue")` [@canaday1996; @sekercioglu2002; @tallamy2021].
`r colorize("Predicted extinctions based on species-level vulnerability point to further dietary shifts, favoring increases in invertivorous species ", "blue")`[@cooke2019].`r colorize("Some systems also show significant shifts in the prevalence of different kinds of foragers in birds, for example loss of arboreal foragers in agricultural systems", "blue")`[@bain2020]`r colorize(", and loss of neotropical understory foragers even in protected areas", "blue")` [@pollock2022].

Here we `r colorize("leverage ongoing efforts to assemble functional trait data and recent computational advances to", "blue")` perform the first multi-taxa, multi-realm assessment of functional diversity and composition change through time.
We focus on mammal and bird species as `r colorize("subsets of the world's biodiversity of particular conservation concern that is heavily impacted by anthropogenic change.", "blue")`While examining trends in plants, invertebrates, and other vertebrate species is of equal interest, trait data for those taxa raise additional challenges such as limited and biased species coverage [@fitzjohn2014], a lack of accepted species-level means, and differences in the types of traits collected.
To ensure comparability across taxa in trait type and data quality we therefore focus on mammals and birds.
`r colorize("We include body mass, dietary, foraging and other behavioral traits", "blue")` that were intentionally selected to be representative of a species' Eltonian niche, thereby summarizing the functional role they play in the community [@wilman2014].

We synthesize thousands of mammal and bird functional diversity time series to assess mean trends across communities.
We evaluate 1) whether or not there is a prevailing type of functional change across communities and if community characteristics explain variation in those trends, 2) hypotheses of change for individual trait axes, and 3) the implications of study-level trends for the ecology of those communities.
To address these goals, we present results for general trends that aggregate across all communities, categorical trends that aggregate across communities with similar characteristics, and study-level trends.

# Material and Methods

## Data

We obtained mammal and bird time series from the BioTIME database, a global repository of high-quality assemblage time series `r colorize("collected from the literature and ongoing monitoring efforts", "blue")`.
`r colorize("Data are structured such that a study comprises all data collected following consistent sampling protocols and may contain multiple time series from separate locations.", "blue")` Time series represent full assemblages rather than populations of single species [@dornelas2018], and all time series include abundance of observed species.
Following best practices for the database [@blowes2019], studies with multiple sample locations were split into individual time series following a standardized spatial scale.
Scale was set by a global grid with cell size determined based on the sample extent of studies with only a single location [see @dornelas2018 for details on how sample extents were defined], with the area of each cell set to one standard deviation away from the mean of the single extent locations.
`r colorize("The resulting cell size for our data was approximately 95 ", "blue")`km^2^.
All samples from a study within a single cell were considered to be a single time series location, and species abundances were combined for all samples.

We used trait data from the Elton Trait Database, which consists of species-level means for traits that represent species' multifaceted role in the community [@wilman2014].
Traits include: body mass, diet, nocturnality, forest foraging strata, pelagic use.
For categorical traits, different category levels are represented as binary or percentage use (Table \ref{tab:traitTable}).

```{r traitTable, echo=FALSE}
tibble(Trait = c(rep("Diet", 10), rep("Foraging Strata", 7), rep("Pelagic Specialist", 2), rep("Nocturnal", 2), rep("Crepuscular", 2), rep("Diurnal", 2), "Body Mass"),
       Category = c("Invertebrate", "Mammals and Birds", "Reptiles", "Fish", "Unknown Vertebrates", "Scavenging", "Fruit", "Nectar", "Seeds", "Other Plant", 
                    "Below water surface", "water surface", "ground", "understory", "> 2m, below canopy", "canopy", "aerial",
                    rep(c("yes", "no"), 4),
                    "-"),
       Taxa = c(rep("Bird and Mammal", 10), rep("Bird", 7), rep("Bird", 2), rep("Bird and Mammal", 2), rep("Mammal", 4), "Bird and Mammal"),
       `Data Type` = c(rep("percentage consumed", 10), rep("percentage of use", 7), rep("binary", 8), "continous, in grams")) %>%
  kable(format = "latex", caption = "Description of the traits included in the analysis broken down by categories at data type.") %>%
  collapse_rows(columns = 1:4) %>%
  column_spec(3:4, width = "3cm") %>%
  kable_paper()
```

In order to ensure taxonomic consistency across datasets, BioTIME species were paired with trait data based on their species identifier from the Integrated Taxonomic Information System database (retrieved 09-15-2020 from the on-line database, <https://doi.org/10.5066/F7KH0KBK>), obtained through the `taxadb` R package [@norman2020; @rcoreteam2021].
If more than one species in the assemblage data resolved to the same identifier, observations were considered the same species.
For trait data, traits for all species of the same identifier were averaged.
Only studies for which at least 75% of species had trait data were included.
In order to have a sufficient number of species to calculate functional diversity metrics, years with fewer than 5 species observed were also excluded.
Sensitivity analyses were conducted for the trait coverage threshold and duration of included time series by rerunning analyses with increasingly conservative cutoffs including 85% and 90 % coverage trait coverage and durations longer than 3, 4, 5, and 10 years.

`r colorize("41 of the 2432 time series included in our dataset had a variable number of samples within years.", "blue")` To account for this inconsistency in sampling effort we used sample-based rarefaction by bootstrap resampling within years for each time series based on the smallest number of samples in a year for that time series.
`r colorize("We made 200 rarefied samples for over sampled years and calculated diversity metrics for each sample, then took the median for each metric as the value for that year and time series. 28 of the affect studies required under sampling by only one sample, and only three time series required under sampling by more than ten samples.", "blue")`

Our final dataset included 2,432 time series from 50 studies in 21 countries and 12 biomes using 6 different traits (Fig \ref{fig:taxaMap}).
Data came from both terrestrial and marine realms and five biomes (Global, Polar/Temperate, Temperate, Temperate/Tropical, Tropical).
The earliest sample was in 1923 and the most recent was in 2014.
`r colorize("While it is not possible with available data to directly assess the level of human impact occurring for each study, we include binary protection status as a coarse indicator of impact level. However, protected areas were almost exclusively from temperate terrestrial studies (with one tropical study), so results are confounded by multiple other study characteristics.", "blue")` For a full breakdown of studies and their characteristics, see the supplement. Our final dataset reflects many of the data biases that make global synthesis work challenging, including geographic bias, a bias away from areas currently under the greatest threat, and a bias towards shorter time series. We address these shortcomings and their potential impact on our results in the discussion.

```{r taxaMap, fig.cap="A) Map of time series locations with points colored by taxa, and B) histograms of time series duration broken down by taxa.", echo=FALSE}
knitr::include_graphics(here::here("figures/study_map_hist.jpeg"))
```

## Diversity Metrics

We calculated yearly metrics of functional and species diversity for each time series.
Species-based metrics include species richness (*S*) and Jaccard similarity (*J*) as a measure of turnover.
Jaccard similarity was calculated relative to the first observed year for a time series.
A negative trend in *J* would therefore indicate `r colorize("decreasing similiarity", "blue")`.
We did not impose a correction for unobserved species as non-parametric estimators do not assign species identities to corrected richness values, and therefore could not be propagated to the functional diversity metrics.

Functional diversity metrics were calculated using the *dbFD* function from the *FD* R package [@laliberté2010].
Here we report functional richness (*FRic*), functional evenness (*FEve*), and functional divergence (*FDiv*) which together describe three complementary characteristics of the functional space [@mason2005; @hillebrand2009].
*FRic* assesses the volume of the trait space occupied by species in the community, with higher values indicating communities with species of more extreme trait values.
*FEve* describes how species are distributed across the trait space and how abundance is distributed across species.
Higher values of *FEve* indicate more even spacing of species in the trait space and individuals across species.
*FDiv* measures the degree to which species and their abundances maximize differences in the functional space.
Higher values of *FDiv* therefore correspond to communities where many highly abundant species are on the edges of the trait space.

We also calculated the community-weighted mean (*CWM*) of all continuous traits to examine `r colorize("changes", "blue")` in the distribution of each trait.
`r colorize("Wholesale shifts in the trait space due to changes of trait means could occur even while the shape of the multidimensional trait space, as defined by functional richness, evenness, and divergence, is maintained. CWM's are therefore a way to assess whether or not turnover is occurring and what the nature of the shift may be. Hereafter we refer to results for functional metrics in two groups: functional diversity metrics (FRic, FEve, FDiv) and composition metrics (trait CWM's).", "blue")`

All available trait data for each study were included in functional diversity calculations with the exception of traits that were the same value for all observed species in the study.
For variables with multiple levels each level was included as a separate trait axis.
Continuous traits were z-score scaled to give each trait equal weight in the trait space [@leps2006; @schleuter2010].
`r colorize("Before calculating diversity metrics, dbFD reduces the dimensionality of the trait space by performing PCoA. We limited the number of included PCoA axes to", "blue")` the maximum number of traits that fulfills the criteria $s >= 2^t$, where $s$ is the number of species and *t* is the number of traits.
This restriction allows for enough axes to capture the trait space while maintaining computational feasibility [@blonder2018].
Metrics incorporated weighting based on species abundance.

## Null Models

To assess functional change independent of species richness we calculated the standardized effect size (SES) for each of the three functional diversity metrics (*FRic*, *FEve*, *FDiv*) from null estimates [@swenson2012].
Null model corrections allow us to assess the degree to which the observed functional diversity metric deviates from the value expected by chance in a randomly assembled community.
Null estimates were calculated for each rarefied sample by randomly sampling species from the species pool for each year and randomly assigning observed abundances to species.
`r colorize("Species pools were unique for each time series and included all species observed over the course of sampling, therefore accounting for geographic restrictions in species availability", "blue")`.
This process was repeated 500 times to get an estimate and standard deviation of the null expectation for the metric for each rarefaction sample for that time series.
We used these values to calculate SES using the following formula: $SES = [F_{obs} - mean_{(F_{null})}]/SD_{(F_{null})}$.
We then calculated the median SES estimate for each metric from all the rarefaction samples for a time series.
SES estimates can be interpreted as how much of the functional characteristic (richness, evenness, divergence) was observed beyond what was expected by chance for a community of that species richness.
`r colorize("This approach will be less accurate for shorter time series, as we likely will not have captured all available species in the true species pool, but it is impossible to know whether the mean estimate from the null model is an over or under estimate without knowing the functional characteristics of the missing species.", "blue")`

## Analysis

We estimated general trends `r colorize("across bird and mammals communities", "blue")` for each diversity metric using a linear mixed effects model with a random slope and intercept for each study and each location nested within the study, `r colorize("methods which deal well with the inherent imbalances in our data.", "blue")` We fit 18 individual *CWM* models, `r colorize("one for each trait included in the analsysis", "blue")`.
All time series with data for a given trait were included in the corresponding *CWM* model.
We estimated study level trends using individual linear models.
For studies with more than one location we fit a random slope and intercept for time series.
Some study-level models could not be fit for five studies for at least one metric due to data limitations, but those studies were still included in the general models.
They represented 13 of 1350 study-level models fit for each metric.
For further details see the supplement.
Where appropriate, response variables were $log$ or $log(x+1)$ transformed to better fit model assumptions of residual normality.

To test for categorical trends within and between different levels of taxa, biome, realm, and protection status we fit separate models with each of those factors added as a predictor interacting with time to the original model structure.
We estimated within-level slopes and calculated between-level contrasts using the *emmeans* package [v1.8.2, @lenth2022].
`r colorize("For some levels of the categorical variables we did not have a sufficient number of studies to estimate a general trend, we therefore only interpret results for levels with more than three studies.", "blue")` We assessed the impact of time series duration and start year on study-level trends using linear models with duration and start year as predictors.
All models in our analysis were fit using the *lme4* (v1.1-30) package in R (v4.2.1) and p-values were calculated by Satterthwaite's degrees of freedom method using the *lmerTest* (v3.1-3) package with a significance level of $\alpha = 0.05$ [@bates2015; @kuznetsova2017; @rcoreteam2023].

To correct for false positives due to multiple testing, we assess significance of the trends for the fixed effect of time based on adjusted p-values.
Adjusted p-values were calculated using the Benjamini-Hochberg procedure, which controls the false discovery rate.
The procedure was implemented by the function `p.adjust` in the R package *stats* (v4.2.1).

# Results

We found no significant general `r colorize("temporal", "blue")` trend in species richness or functional diversity metrics `r colorize("including functional richness, evenness, or divergence", "blue")` (observed or `r colorize("corrected", "blue")`) (Fig \ref{fig:timeseriesPlot}).
We did find a significant overall decrease in Jaccard similarity, indicating accumulating changes in species composition.
Non-significant overall `r colorize("temporal", "blue")` trends indicate that although some studies experience increasing or decreasing trends, the average trend across studies `r colorize("was not significantly different from zero", "blue")` (Table \ref{tab:resultsTab}).
`r colorize("Categorical", "blue")` trends for different taxa, biomes, realms, or protection statuses were also non-significant for species richness and functional diversity metrics.

```{r timeseriesPlot, fig.cap="Plots of time series-level trends with line color corresponding to climatic region, with data points in grey and the overall mean slope for a metric in black for A) log species richness, B) Jaccard similarity, C) Functional Richness SES, D) Functional Divergence SES, and E) Functional Evenness SES", echo=FALSE}
knitr::include_graphics(here::here("figures/3met_long.jpeg"))
```

```{r resultsTab, echo=FALSE}
read_csv(here::here("figures/man_model_table.csv")) %>%
  select(metric, everything(), -REML, -bobyqa) %>%
  select(-c(n_ts, formula, lower.CL, upper.CL, significant)) %>%
  filter(group != "Residual" | is.na(group)) %>%
  mutate(metric = ifelse(metric == "Jaccard_base", "log(Jaccard + 1)", metric),
         metric = ifelse(metric == "S", "log(Species Richness)", metric),
         effect = ifelse(effect == "ran_pars", "random", effect),
         group = case_when(
           group == "rarefyID:study_id" ~ "time series within study",
           group == "study_id" ~ "study",
           group == "Residual" ~ "residual",
           TRUE ~ group),
         term = case_when(
           term == "(Intercept)" ~ "Intercept",
           term == "year_scaled" ~ "Year",
           term == "sd__(Intercept)" ~ "SD Intercept",
           term == "cor__(Intercept).year_scaled" ~ "Corr(Intercept, Year)",
           term == "sd__year_scaled" ~ "SD Year",
           term == "sd__Observation" ~ "SD Observation",
           TRUE ~ term
         )) %>%
  mutate(across(c(estimate, std.error, p.value), round, 2),
         p.value = ifelse(p.value == 0, "<0.001", p.value)) %>%
  arrange(metric != "Species Richness", metric, !is.na(group), group, term == "Corr(Intercept, Year)", term) %>%
  dplyr::rename(grouping = group) %>%
  kable(format = "latex", caption = "Model estimates and statistics for general trend models for species richness, Jaccard similarity, and corrected functional diversity metrics. Additional model estimates, including CWM models, can be found in the supplement.") %>%
  column_spec(3, width = "2cm") %>%
  collapse_rows(columns = 1:3, valign = "middle") %>%
  kable_paper() %>%
  kable_styling(latex_options= c("scale_down", "hold_position")) #%>%
  #save_kable(here::here("figures/tables/results_table.jpeg"))
```

We did find significant categorical trends for Jaccard similarity and two dietary *CWM's*.
While Jaccard similarity was decreasing in the general trend and there were significant categorical slopes for birds, terrestrial, and temperate studies, there was no significant slope for marine, mammal, or tropical studies.
We also found a significant decrease in Jaccard similarity in unprotected areas only, with no trend for protected areas.
We found significant dietary shifts across communities, with a significant general trend of increasing fish consumption, which was also reflected in increasing fish consumption trends for bird studies and temperate studies.
We found significantly decreasing trend for vertebrate consumption in Marine studies and Tropical studies.
The trend for marine vertebrate consumption was significantly more negative than for terrestrial studies and the temperate trend was significantly more negative than tropical studies.

At the study level, 11 of 50 studies exhibited a significant trend in species richness and 11 exhibited significant turnover.
For uncorrected functional diversity metrics, 11 of 50 studies exhibited a trend in a least one metric, and 7 of 50 studies exhibited a significant trend for at least one `r colorize("corrected", "blue")` metric (Table \ref{tab:trendTab}).
In general, there were more significant trends for uncorrected metrics, with some disappearing after correction, indicating that those trends were likely due to changes in the number of species.
Study-level slopes were not significantly related to start year or duration of time series for any metric.

```{r trendTab, echo = FALSE}
library(tidyverse)

load(here::here("data/model_output.rda"))

model_output %>%
  filter(term == "year_scaled", p.adjust < 0.05) %>%
  filter(metric %in% c("SES_FRic", "SES_FEve", "SES_FDiv", "Jaccard_base", "S", "FRic", "FEve", "FDiv")) %>%
  mutate(positive = ifelse(estimate > 0, "+", "-")) %>%
  group_by(metric, positive) %>%
  summarise(n_study_id = n_distinct(study_id)) %>%
  pivot_wider(names_from = "metric", values_from = "n_study_id", values_fill = 0) %>%
  select(positive, S, Jaccard_base, FRic, FEve, FDiv, SES_FRic, SES_FEve, SES_FDiv) %>%
  kable(format = "latex", 
        #col.names = c("", "S", "Jaccard Similarity", "FRic", "FEve", "FDiv", "SES FRic", "SES FEve", "SES FDiv"),
        col.names = c("", "richness", "Jaccard Similarity", "richness", "evenness", "divergence", "SES richness", "SES evenness", "SES divergence"),
        caption = "Number of studies that experienced a significant trend in each calculated metric out of 50 total studies. Both uncorrected and corrected (SES) versions of functional metrics are included.", 
        booktabs = TRUE, align = "c", escape = FALSE) %>%
  kableExtra::add_header_above(c("", "species" = 2, "functional" = 6)) %>%
  kable_styling("striped") #%>%
  #save_kable(here::here("figures/tables/trend_table.jpeg"))

```

We assessed the sensitivity of general trend results to major data processing decision by rerunning models with increasingly conservative subsets of the data.
Trends for Jaccard similarity and fish consumption were not sensitive to either time series duration or trait coverage.
After excluding time series with less than three years we found an increasing trend for body mass that remained after excluding time series of less than four and five years.
The body size trend did however disappear after excluding time series of less then 10 years.
A complete list of models run in the sensitivity analysis and their results can be found in the supplement.

# Discussion

Our study represents the largest broad-scale multi-taxa assessment of functional change through time to date, giving a first look at general, categorical, and study-level local trends in functional diversity in mammal and bird communities.
Our work complements efforts on a similar scale assessing temporal change in species and phyllogenetic diversity across taxa [@daru2021; @dornelas2014; @li2020; @vellend2013l; @vellend2017] or functional diversity change for a single taxa [@barnagaud2017; @jarzyna2016; @trindade-santos2020].
`r colorize("We did not detect an overall trend in any functional diversity metric, corrected or uncorrected. As with previous species-based syntheses, we also found no overall trend in species richness accompanied by increasing dissimilarity through time", "blue")` [@dornelas2018]`r colorize(", indicating that non-significant trends in functional metrics are consistent with similar well-documented species derived trends.", "blue")`

`r colorize("Categorical trends in turnover highlight the types of communities experiencing change in community composition. For example, we found significant turnover for most biomes, realms, and taxa with the exception of marine, tropical, and mammal studies. The lack of turnover in marine studies stands in contrast to other global estimates of biodiversity change that found higher turnover in marine systems than terrestrial", "blue")` [@blowes2019].
`r colorize("However, previous global estimates are dominated by fish communities which we exclude here and are likely driving the overall trend while disguising relative stasis in marine bird and mammal communities. We also found significant turnover for unprotected areas and no significant turnover in protected areas, indicating that in general unprotected communities are experiencing bigger shifts in community composition through time.", "blue")`

`r colorize("Surprisingly, we found no evidence for some of the strongest predictions for anthropogenic impact. For example, we found no trend for mean body mass in bird or mammal communities. In fact, sensitivity results indicated that after excluding the shortest time series there is evidence of an increasing trend in body size. This result could be an indication that either body size is not changing significantly due to climate change, opposing pressures such as urbanization or carnivore restablishment are overshadowing climate change impacts, or current shifts are happening at an intraspecific level not captured by our data. Additionally, the majority of the studies in our dataset draw from areas that may have experienced significant loss of large-bodied species before the observation window, with contemporary loss rates slowing", "blue")` [@fritz2009].
`r colorize("Trends could be significantly different for the same time periods in regions of sub-Saharan Africa, for example, which has poor representation in our dataset but where megafauna exist on the landscape and are increasingly threatened", "blue")` [@ripple2015].

`r colorize("We similarly did not find evidence of changes in trait distributions predicted by previous global change work. We found no changes in the prevalence of different foraging strategies, despite documented losses of understory birds in the neotropics and some evidence that agricultural incursion particularly threatens arboreal species. Those shifts may therefore be the result of specific contexts and not generalizable to bird communities across the globe. Rather than declines in insectivory and increases in invertivory predicted by changes in resource availability, we instead found changes in the degree of consumption of fish and vertebrates. Neither increasing trends for fish consumption nor decreasing trends for vertebrate consumption in marine and tropical studies are not obviously consistent with known changes in those contexts and warrant further examination.", "blue")`

## Study-level functional change

```{r studyChange, echo = FALSE}

load(here::here("data/change_group_ses_desc.rda"))

change_group_ses_desc[change_group_ses_desc == "pos"] <- "+"
change_group_ses_desc[change_group_ses_desc == "neg"] <- "-"

change_group_ses_desc %>%
  select(type = change_description, everything()) %>%
  mutate(n = as.character(n),
           n = ifelse(n == "7", "32", n)) %>%
  mutate(type = ifelse(n == "32", "no change", type)) %>%
  rename(`Study Count` = `n`) %>%
     kable(format = "latex",
        caption = "Table giving the count of number of studies for different kinds of functional change, broken down into no change functional change without species richness change, loss of redundancy, and increases in species richness. Trends direction is indicate by a positive or negative sign, with a 0 indicating no trend.", align = "c",
        latex_options= c("scale_down", "hold_position")) %>% 
  collapse_rows(columns = c(1,2)) %>%
  column_spec(1, width = "4cm")

```

`r colorize("While trends aggregating across studies did not indicate one prevailing kind of functional change, post hoc examination of trends for individual studies revealed four distinct change groups. These groups are categorized based on the concurrent trends in species and functional metrics and can be broadly described as: no change in species or functional diversity metrics, functional change independent of species richness, loss of redundancy, and increase in species richness", "blue")` (Table \ref{tab:studyChange}).
`r colorize("The majority of studies (32 studies) exhibited no trend in any species or functional diversity metric. Contrary to the expectation due to anthropogenic and global change stressors, these communities do not show significant changes over the course of the observation window. Studies in this group span the distribution of study durations excluding only the very longest running studies, with the longest no change time series lasting 33 years. They also included both bird and mammal studies and only seven were located in protected areas, indicating that the lack of trend is not restricted to a specific ecological context or those communities most insulated from human impact.", "blue")`

`r colorize("The lack of trend could be the result of multiple possible scenarios. First, these may be communities resisting perturbations or simply not experiencing significant perturbations. Given the studies in this group come from all possible taxa, realms, biomes, and protection statuses, evidence points to communities resisting perturbation. Alternatively, these may be communities that have experienced or continue to experience significant stress, but lost species or functional diversity outside the observation window. This could be true particularly for North American mammal communities where trophic downgrading and megafaunal losses occurred hundreds of years ago ", "blue")`[@estes2011].
`r colorize("Third, these communities may be experiencing directional shifts undetectable by available data. For example, species-level trait data do not capture intraspecific shifts in the trait space, which can have significant implications for ecological processes.", "blue")`

`r colorize("The next largest group of studies (7 studies) were those with no species richness trend that showed a trend in at least one functional diversity metric. Based on a species approach alone, these communities apear to either be experiencing no change or some change in species identity, with unknown implications for conservation concern or ecological process. However, with the addition of a functional perspective we found multiple kinds of underlying functional change including increases and decreases in corrected functional richness and divergence and loss of corrected functional evenness. We therefore found evidence of both increases and decreases in the functional capacity of communities independent of species-based metrics, reinforcing the importance of moving beyond soley species-based approaches for biodiversity assessment", "blue")` [@hillebrand2017; @santini2017].

`r colorize("9 studies showed loss of functional redundancy. These studies are characterized by a loss of species richness with no shift in functional diversity metrics, indicating the lost species were functionally similar enough to another species to not impact the overall functional structure. Redundancy is critical for maintaining ecological processes in the face of disturbance as it ensures  functional roles are maintained", "blue")` [@biggs2020]`r colorize(", and a loss of functional redundancy likeley reduces a community's capacity to respond to future stressors ", "blue")`[@reich2012]`r colorize(". These communities are actually fairing better than expected looking at species-based metrics alone, but are also in a precarious position for maintaining ecological function into the future", "blue")` [@pimiento2020a].

`r colorize("The final two studies showed an increase in species richness and uncorrected functional richness, with only one showing an increase in functional richness beyond what was expected by chance (corrected functional richness). While this can be an indication of increasing functional capacity or reestabilishment of previously lost species, it can also be the result of less desirable process like the introduction of invasive species. Notably, both of these studies were from temperate, terrestrial bird communities in Scandinavian countries with a unique environment and conservation context. Still, results are consistent with other broad-scale studies of bird communities, which found loss of common, functionally general species alongside increases in rare species ", "blue")` [@inger2015; @rosenberg2019; @schipper2016].

## Conservation Implications

`r colorize("While we found no over all trends in functional metrics, our results should not be interpreted as an indication that the ongoing biodiversity crisis is less severe than previously described, or that there is no concern for functional change as a result of anthropogenic impact. In fact, study-level trends indicate quite the opposite, that functional shifts with negative or yet unknown implications for ecosystem processes may be going undetected by common species-based approaches. For example, loss of evenness in communities with constant species richness may be an first sign of a community being impacted by environmental change, with negative implications for stability and function.", "blue")`

`r colorize("One of the biggest threats to biodiversity is the wholesale conversion of natural areas to urban or human-dominated landscapes ", "blue")`[@tilman2017].
`r colorize("Typical long term monitoring such as those included in our study stops before this conversion occurs, leaving the resultant precipitous declines in biodiversity unrecorded", "blue")` [@cardinale2014].
`r colorize("This is a known issue with the culture of long term monitoring, and our results should not be removed from that context. Rather, this studies captures communities that are likely experiencing a degree of human intervention, but are still largely nature-dominated.", "blue")`

# Acknowledgments

We thank the scientists who contributed to and maintain the biodiversity databases included in this study, including BioTIME, Amphibio, and Elton Traits.

We thank Dr Shane Blowes and Dr Sarah Supp, whose code we adapted for initial data processing of the BioTIME database.

# References