### 2024-11-21 analysis/exporatory_scripts/mantel_test.Rmd
First commit

---
title: "Mantel Test"
output: html_document
---

```{r}
library(dplyr)
```

get rarefied samples
```{r}
#helper function to load file
loadRData <- function(file_name){
  load(file_name)
  get(ls()[ls() != "file_name"])
}

load(here::here("data/model_data.rda"))
long_duration <- model_data %>% select(rarefyID, duration) %>%
  filter(duration > 2) %>%
  pull(rarefyID)

#get list of rarefaction sample files
path <- here::here("data", "rarefied_samples")
files <- dir(path, "*.rda") %>%
  paste0(path, "/", .)

file_df <- data.frame(file_path = files) %>%
  mutate(rarefyID = str_match(files,"_cell\\s*(.*?)\\s*_sample")[,2]) %>%
  filter(rarefyID %in% long_duration)

get_dist <- function(files){

    rare_comm <- loadRData(files[1])

    # meta <- rare_comm %>%
    #   select(rarefyID, cell, rarefy_resamp, type) %>%
    #   distinct()

    rare_comm <- rare_comm %>%
      column_to_rownames("YEAR") %>%
      dplyr::select(-rarefyID, -cell, -rarefy_resamp, -type)

    mat <- vegdist(rare_comm, method='jaccard', binary=TRUE)

    return(mat)
  }

rarefyid_distmat <- unique(file_df$rarefyID) %>%
  purrr::set_names() %>% 
  purrr::map(function(id){

  #browser()

  files <- file_df %>% filter(rarefyID == id) %>%
    pull(file_path)

  if (length(files) > 1) {

  mat_list <- purrr::map(files, get_dist)
  dist_mat <- Reduce("+", mat_list) / length(mat_list)

  } else {

    dist_mat <- get_dist(files)

  }
  return(dist_mat)
})

dim_check <- purrr::map(rarefyid_distmat, ~ dim(.x)[1] != 2)
```

```{r}
# remove time series with only two time points, won't be able to permute
sub_rarefyid_distmat <- rarefyid_distmat[unlist(dim_check)]

get_mantel <- function(jacc_dist_mat){
  years <- as.integer(names(jacc_dist_mat))
  names(years) <- names(jacc_dist_mat)
  year_dist <- dist(years)

  mantel_out <- mantel(jacc_dist_mat, year_dist, method = "spearman", permutations = 9999)

  return(data.frame(p.value = mantel_out$signif, perm_num = mantel_out$permutations))
}

 mantel_list <- #names(sub_rarefyid_distmat) %>%
#   purrr::set_names() %>% 
  purrr::map(sub_rarefyid_distmat, get_mantel)

 names(mantel_list) <- names(sub_rarefyid_distmat)

 mantel_df <- bind_rows(mantel_list, .id = "rarefyID") %>%
   mutate(p.adjust =  p.adjust(mantel_df$p.value, method = "BH"))

 usethis::use_data(mantel_df)
```